---
title: "Additional coverage plots"
author: "Briana Mittleman"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to make some additional plots of the coverage of 3' seq at locations. I will use the deeptools plots.

For these I will make seperate total and nuclear plots 
```{r}
library(tidyverse)
library(workflowr)
```


###PacBio  

I will use the PAS called using LCL pacbio data. Ankeeta created the file. 
```{bash,eval=F}
mkdir ../data/pacbio
cp /project2/yangili1/ankeetashah/APA_tools/pacbio_3/PACBIO.ANNOTATED.USAGE.BED ../data/pacbio/PACBIO.Annotated.usage.PublicData.bed

#remove chr and sort

sed 's/^chr//' ../data/pacbio/PACBIO.Annotated.usage.PublicData.bed | sort -k1,1 -k2,2n > ../data/pacbio/pacbioPACBIO.Annotated.usage.PublicData.sort.bed

../data/pacbio/pacbioPACBIO.Annotated.usage.PublicData.sort.bed

```

```{bash,eval=F}
sbatch pacbioDT.sh
```

Subset to those falling in intronic regions. I can do this by intersecting with my original annotation file.  
/project2/gilad/briana/genome_anotation_data/RefSeq_annotations/ncbiRefSeq_FormatedallAnnotation.sort.bed

```{bash,eval=F}
sbatch annotatePacBioPASregion.sh
```

I need to parse this like I did my original file.  

```{r}
pacbio=read.table("../data/pacbio/pacbioPACBIO.Annotated.usage.PublicData.sort_withAnnotationparsed.SAF", header = T,stringsAsFactors = F) %>% separate(GeneID, into =c("PASnum", "chr", "start","end", "strand","geneID"),sep=":") %>% separate(geneID, into = c("gene", "loc"), sep="_") %>% filter(loc=="intron") %>% mutate(pASid=paste(PASnum, gene, sep=":"), score=".") %>% select(Chr, Start, End, pASid, score, strand)

write.table(pacbio, file="../data/pacbio/IntronicPacBioPAS.bed", col.names = F, row.names = F, quote = F, sep="\t")
```
```{bash,eval=F}
sbatch pacbioIntronicDT.sh
```


###intronic PAS  

```{bash,eval=F}
sbatch IntronicPASDT.sh

python chooseAnno2PAS_pacbio.py ../data/pacbio/pacbioPACBIO.Annotated.usage.PublicData.sort_withAnnotation.bed ../data/pacbio/pacbioPACBIO.Annotated.usage.PublicData.sort_withAnnotationparsed.SAF


```



