---
title: "Check for first intron enrichment"
author: "Briana Mittleman"
date: "11/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Check for enrichment in first intron.  

This uses files created in the verifyFirstintronresult analysis. 

```{r}
library(workflowr)
library(tidyverse)
library(reshape2)
```

```{r}
pas2intron= read.table( "../data/intron_analysis/NuclearIntronPASwithWhichintron.txt", stringsAsFactors = F, header = T)
```


I need to know how many 1st introns there are, how many second introns there are... 

The number of introns in these genes is in the nintron column. 

I can group the number of genes with each number of introns. 



```{r}
pas2intron_genes= pas2intron %>% select(gene, nintron) %>% unique() %>% group_by(nintron) %>% summarize(nGenes=n())

head(pas2intron_genes)


```

I want to add these numbers but only include below. For example, all of these have 1 intron, only those above 5 have 5 introns. 

Do this manually for 1-10 first. 


```{r}
n1intron=sum(pas2intron_genes$nGenes)
n2intron=sum(pas2intron_genes$nGenes)
n3intron=sum(pas2intron_genes$nGenes)
n4intron= pas2intron_genes %>% filter(nintron >=4) 
n5intron= pas2intron_genes %>% filter(nintron >=5) 
n6intron= pas2intron_genes %>% filter(nintron >=6) 
n7intron= pas2intron_genes %>% filter(nintron >=7) 
n8intron= pas2intron_genes %>% filter(nintron >=8) 
n9intron= pas2intron_genes %>% filter(nintron >=9) 
n10intron= pas2intron_genes %>% filter(nintron >=10) 


nGenesum=c(n1intron,n2intron,n3intron, sum(n4intron$nGenes),sum(n5intron$nGenes),sum(n6intron$nGenes),sum(n7intron$nGenes),sum(n8intron$nGenes),sum(n9intron$nGenes),sum(n10intron$nGenes))
```


```{r}
pas2intron_intron_grouped=pas2intron %>% group_by(Intronid) %>% summarise(nBin=n(), meanSize=mean(intronLength))  %>% mutate(normNBin=nBin/meanSize) 

pas2intron_intron_grouped_small=pas2intron_intron_grouped %>% filter(Intronid <=10) 

pas2intron_intron_grouped_small_withnintron=as.data.frame(cbind(pas2intron_intron_grouped_small, nGenesum)) %>% mutate(NormbylengthandN=normNBin/nGenesum)

ggplot(pas2intron_intron_grouped_small_withnintron, aes(x=Intronid, y=NormbylengthandN)) +geom_bar(stat="identity") + labs(title="PAS by Intron \n normalized by intron mean length and number of introns tested", y="normalized number in intron category", x="intron category")



```


Look at just genes with 3 introns.  

```{r}
pas2intron_3introns=pas2intron %>% filter(nintron==3) %>% group_by(Intronid) %>% summarise(nBin=n(), meanSize=mean(intronLength))  %>% mutate(normNBin=nBin/meanSize)

ggplot(pas2intron_3introns,aes(y=normNBin,x=Intronid)) + geom_bar(stat = "identity") + labs(x="Intron in Gene",y="Number introns / average intron length", title="Intron position for genes with 3 introns")
```

