---
title: "Compare to annotated PAS"
author: "Briana Mittleman"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(workflowr)
library(tidyverse)
library(reshape2)
```


I will se the annotated PAS from the Tian lab database (http://exon.umdnj.edu/polya_db/v3/misc/download.php)

```{bash,eval=F}
mkdir ../data/AnnotatedPAS/
  
#file =human.PAS.txt
```

I want to make this into a file I can overlap with my PAS. In order to know what resolution I should use for calling a PAS the same, I will look for the closest annotated PAS to each of my sites. To do this I will need to create a bed file with these.  


```{bash,eval=F}
python annotatedPAS2bed.py
sort -k1,1 -k2,2n ../data/AnnotatedPAS/human.PAS.bed > ../data/AnnotatedPAS/human.PAS.sort.bed

sbatch closestannotated.sh
```


```{r}
dist=read.table("../data/AnnotatedPAS/DistanceMyPAS2Anno.bed", col.names = c("chr", "start","end","myPAS", "score","strand","chr2", "start2", "end2", "anno", "score2", "strand2", "distance"),stringsAsFactors = F)
```

Plot the distance.

```{r}
ggplot(dist,aes(x=distance))+ geom_histogram(bins=300) + xlim(-25, 25)
```

Looks like about 10 basepairs is ok resolution. I need to make sure these map 1 to 1 when you filter these.  

```{r}
PAS_withmatch=dist %>% filter(abs(distance)<=10) %>% select(myPAS,anno) %>% unique() %>% separate(myPAS, into=c("pasNum", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc"), sep="_")

ggplot(PAS_withmatch,aes(x=loc)) + geom_histogram(stat="count")
```

I want to look at those I find that they do not.

```{r}
allMyPAS=read.table("../data/PAS/APAPAS_GeneLocAnno.5perc.sort.bed",stringsAsFactors = F, col.names = c("chr","start","end", "PASID", "score","strand"))  %>% separate(PASID, into=c("pasNum", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc"), sep="_") %>% mutate(withAnno=ifelse(pasNum %in% PAS_withmatch$pasNum, "Yes","No"))

PASnoMatch=allMyPAS %>% anti_join(PAS_withmatch,by="pasNum")
```

```{r}
ggplot(allMyPAS,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```


Look at total and nuclear seperatly. 

```{bash,eval=F}
python NuclearPAS_5per.bed.py


python TotalPAS_5perc.bed.py
sort -k1,1 -k2,2n ../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.bed > ../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.sort.bed

sort -k1,1 -k2,2n ../data/PAS/APApeak_Peaks_GeneLocAnno.Total.5perc.bed > ../data/PAS/APApeak_Peaks_GeneLocAnno.Total.5perc.sort.bed

```
Run the distance script with these.  

```{bash,eval=F}
sbatch closestannotated_byfrac.sh
```

###Total
```{r}
Totaldist=read.table("../data/AnnotatedPAS/Total_DistanceMyPAS2Anno.bed", col.names = c("chr", "start","end","myPAS", "score","strand","chr2", "start2", "end2", "anno", "score2", "strand2", "distance"),stringsAsFactors = F)

ggplot(Totaldist,aes(x=distance))+ geom_histogram(bins=300) + xlim(-25, 25)

```
```{r}
Totaldist_withAnno=Totaldist %>% filter(abs(distance)<=10) %>% select(myPAS,anno) %>% unique() %>% separate(myPAS, into=c("pasNum", "geneID", "loc"), sep=":")

allTotalPAS=read.table("../data/PAS/APApeak_Peaks_GeneLocAnno.Total.5perc.sort.bed",stringsAsFactors = F, col.names = c("chr","start","end", "PASID", "score","strand"))  %>% separate(PASID, into=c("pasNum", "geneID", "loc"), sep=":") %>% mutate(withAnno=ifelse(pasNum %in% Totaldist_withAnno$pasNum, "Yes","No"))
```

```{r}
ggplot(allTotalPAS,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "TotalPAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

###Nuclear  
```{r}
Nucleardist=read.table("../data/AnnotatedPAS/Nuclear_DistanceMyPAS2Anno.bed", col.names = c("chr", "start","end","myPAS", "score","strand","chr2", "start2", "end2", "anno", "score2", "strand2", "distance"),stringsAsFactors = F)

ggplot(Nucleardist,aes(x=distance))+ geom_histogram(bins=300) + xlim(-25, 25)

```
```{r}
Nucleardist_withAnno=Nucleardist %>% filter(abs(distance)<=10) %>% select(myPAS,anno) %>% unique() %>% separate(myPAS, into=c("pasNum", "geneID", "loc"), sep=":")

allNuclearPAS=read.table("../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.sort.bed",stringsAsFactors = F, col.names = c("chr","start","end", "PASID", "score","strand"))  %>% separate(PASID, into=c("pasNum", "geneID", "loc"), sep=":") %>% mutate(withAnno=ifelse(pasNum %in% Nucleardist_withAnno$pasNum, "Yes","No"))
```

```{r}
ggplot(allNuclearPAS,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

Nuclear specific:

```{r}
NuclearSpec=allNuclearPAS %>% anti_join(allTotalPAS,by = "pasNum")

ggplot(NuclearSpec,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear  Specific PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

Total Specific:
```{r}
TotalSpec=allTotalPAS %>% anti_join(allNuclearPAS,by = "pasNum")

ggplot(TotalSpec,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Total  Specific PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

##By usage

```{r}
NuclearMeanUsage=read.table("../data/PAS/NuclearPASMeanUsage.txt",header = T, stringsAsFactors = F) %>% separate(ID, into=c("chr", "start", "end", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc", "strand", "PAS"),sep="_")

TotalMeanUsage=read.table("../data/PAS/TotalPASMeanUsage.txt",header = T, stringsAsFactors = F) %>% separate(ID, into=c("chr", "start", "end", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc", "strand", "PAS"),sep="_")

NuclearMeanUsage_25= NuclearMeanUsage %>% filter(meanUsage >.25)


allNuclearPAS_25= allNuclearPAS %>% mutate(PAS=paste("peak", pasNum,sep="")) %>% anti_join(NuclearMeanUsage_25, by="PAS")

ggplot(allNuclearPAS_25,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear PAS 25% by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")



NuclearMeanUsage_50= NuclearMeanUsage %>% filter(meanUsage >.5)


allNuclearPAS_50= allNuclearPAS %>% mutate(PAS=paste("peak", pasNum,sep="")) %>% anti_join(NuclearMeanUsage_50, by="PAS")

ggplot(allNuclearPAS_50,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear PAS 50% by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2") 


NuclearMeanUsage_75= NuclearMeanUsage %>% filter(meanUsage >.75)


allNuclearPAS_75= allNuclearPAS %>% mutate(PAS=paste("peak", pasNum,sep="")) %>% anti_join(NuclearMeanUsage_75, by="PAS")

ggplot(allNuclearPAS_75,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear PAS 75% by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2") 
```
Proportion of previosly identified:

```{r}
allNuclearPAS %>% group_by(withAnno) %>% summarise(All=n()) %>% ungroup() %>% mutate(Prop=All/sum(All))

allNuclearPAS_25 %>% group_by(withAnno) %>% summarise(TwentyFive=n()) %>% ungroup() %>% mutate(Prop=TwentyFive/sum(TwentyFive))


allNuclearPAS_50 %>% group_by(withAnno) %>% summarise(Fifty=n()) %>% ungroup() %>% mutate(Prop=Fifty/sum(Fifty))

allNuclearPAS_75 %>% group_by(withAnno) %>% summarise(Seventyfive=n()) %>% ungroup() %>% mutate(Prop=Seventyfive/sum(Seventyfive))
```
Do this for usage 1-100%


```{r}


withAnno_nuc=function(fraction){
  NuclearMeanUsage_prop= NuclearMeanUsage %>% filter(meanUsage >= fraction)
  allNuclearPAS_prop= allNuclearPAS %>% mutate(PAS=paste("peak", pasNum,sep="")) %>% anti_join(NuclearMeanUsage_prop, by="PAS") %>% group_by(withAnno) %>% summarise(All=n()) %>% ungroup() %>% mutate(Prop=All/sum(All)) %>% filter(withAnno=="Yes")
  #print(paste(fraction,allNuclearPAS_prop$Prop))

return(allNuclearPAS_prop$Prop)
}

propYes=c()

cutoffs=seq(from=.1, to=1, by=.05)
for (val in cutoffs){
  newVal=withAnno_nuc(val)
  propYes=c(propYes,newVal )
}
nucCuttoff=cbind(cutoff=cutoffs,Nuclear=propYes)


withAnno_tot=function(fraction){
  TotalMeanUsage_prop=TotalMeanUsage %>% filter(meanUsage >= fraction)
  allTotalPAS_prop= allTotalPAS %>% mutate(PAS=paste("peak", pasNum,sep="")) %>% anti_join(TotalMeanUsage_prop, by="PAS") %>% group_by(withAnno) %>% summarise(All=n()) %>% ungroup() %>% mutate(Prop=All/sum(All)) %>% filter(withAnno=="Yes")
  #print(paste(fraction,allNuclearPAS_prop$Prop))
return(allTotalPAS_prop$Prop)
}

propYesTot=c()

cutoffs=seq(from=.1, to=1, by=.05)
for (val in cutoffs){
  newVal=withAnno_tot(val)
  propYesTot=c(propYesTot,newVal )
}
AllCuttoff=as.data.frame(cbind(cutoff=cutoffs,Nuclear=propYes, Total=propYesTot))


```


```{r}
AllCuttoff_melt=melt(AllCuttoff,id.vars="cutoff", variable.name = "Fraction", value.name = "PropwithAnno")


ggplot(AllCuttoff_melt, aes(x=cutoff, col=Fraction, y= PropwithAnno)) + geom_line(size=2) + scale_color_brewer(palette = "Dark2") + labs(title="Cumulative proportion of PAS in annoation \n by usage filter", y="Proportion in Annotation", x="Usage Filter")
```


