---
title: "Compare to annotated PAS"
author: "Briana Mittleman"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(workflowr)
library(tidyverse)

```


I will se the annotated PAS from the Tian lab database (http://exon.umdnj.edu/polya_db/v3/misc/download.php)

```{bash,eval=F}
mkdir ../data/AnnotatedPAS/
  
#file =human.PAS.txt
```

I want to make this into a file I can overlap with my PAS. In order to know what resolution I should use for calling a PAS the same, I will look for the closest annotated PAS to each of my sites. To do this I will need to create a bed file with these.  


```{bash,eval=F}
python annotatedPAS2bed.py
sort -k1,1 -k2,2n ../data/AnnotatedPAS/human.PAS.bed > ../data/AnnotatedPAS/human.PAS.sort.bed

sbatch closestannotated.sh
```


```{r}
dist=read.table("../data/AnnotatedPAS/DistanceMyPAS2Anno.bed", col.names = c("chr", "start","end","myPAS", "score","strand","chr2", "start2", "end2", "anno", "score2", "strand2", "distance"),stringsAsFactors = F)
```

Plot the distance.

```{r}
ggplot(dist,aes(x=distance))+ geom_histogram(bins=300) + xlim(-25, 25)
```

Looks like about 10 basepairs is ok resolution. I need to make sure these map 1 to 1 when you filter these.  

```{r}
PAS_withmatch=dist %>% filter(abs(distance)<=10) %>% select(myPAS,anno) %>% unique() %>% separate(myPAS, into=c("pasNum", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc"), sep="_")

ggplot(PAS_withmatch,aes(x=loc)) + geom_histogram(stat="count")
```

I want to look at those I find that they do not.

```{r}
allMyPAS=read.table("../data/PAS/APAPAS_GeneLocAnno.5perc.sort.bed",stringsAsFactors = F, col.names = c("chr","start","end", "PASID", "score","strand"))  %>% separate(PASID, into=c("pasNum", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc"), sep="_") %>% mutate(withAnno=ifelse(pasNum %in% PAS_withmatch$pasNum, "Yes","No"))

PASnoMatch=allMyPAS %>% anti_join(PAS_withmatch,by="pasNum")
```

```{r}
ggplot(allMyPAS,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```


Look at total and nuclear seperatly. 

```{bash,eval=F}
python NuclearPAS_5per.bed.py


python TotalPAS_5perc.bed.py
sort -k1,1 -k2,2n ../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.bed > ../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.sort.bed

sort -k1,1 -k2,2n ../data/PAS/APApeak_Peaks_GeneLocAnno.Total.5perc.bed > ../data/PAS/APApeak_Peaks_GeneLocAnno.Total.5perc.sort.bed

```
Run the distance script with these.  

```{bash,eval=F}
sbatch closestannotated_byfrac.sh
```

###Total
```{r}
Totaldist=read.table("../data/AnnotatedPAS/Total_DistanceMyPAS2Anno.bed", col.names = c("chr", "start","end","myPAS", "score","strand","chr2", "start2", "end2", "anno", "score2", "strand2", "distance"),stringsAsFactors = F)

ggplot(Totaldist,aes(x=distance))+ geom_histogram(bins=300) + xlim(-25, 25)

```
```{r}
Totaldist_withAnno=Totaldist %>% filter(abs(distance)<=10) %>% select(myPAS,anno) %>% unique() %>% separate(myPAS, into=c("pasNum", "geneID", "loc"), sep=":")

allTotalPAS=read.table("../data/PAS/APApeak_Peaks_GeneLocAnno.Total.5perc.sort.bed",stringsAsFactors = F, col.names = c("chr","start","end", "PASID", "score","strand"))  %>% separate(PASID, into=c("pasNum", "geneID", "loc"), sep=":") %>% mutate(withAnno=ifelse(pasNum %in% Totaldist_withAnno$pasNum, "Yes","No"))
```

```{r}
ggplot(allTotalPAS,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "TotalPAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

###Nuclear  
```{r}
Nucleardist=read.table("../data/AnnotatedPAS/Nuclear_DistanceMyPAS2Anno.bed", col.names = c("chr", "start","end","myPAS", "score","strand","chr2", "start2", "end2", "anno", "score2", "strand2", "distance"),stringsAsFactors = F)

ggplot(Nucleardist,aes(x=distance))+ geom_histogram(bins=300) + xlim(-25, 25)

```
```{r}
Nucleardist_withAnno=Nucleardist %>% filter(abs(distance)<=10) %>% select(myPAS,anno) %>% unique() %>% separate(myPAS, into=c("pasNum", "geneID", "loc"), sep=":")

allNuclearPAS=read.table("../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.sort.bed",stringsAsFactors = F, col.names = c("chr","start","end", "PASID", "score","strand"))  %>% separate(PASID, into=c("pasNum", "geneID", "loc"), sep=":") %>% mutate(withAnno=ifelse(pasNum %in% Nucleardist_withAnno$pasNum, "Yes","No"))
```

```{r}
ggplot(allNuclearPAS,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

Nuclear specific:

```{r}
NuclearSpec=allNuclearPAS %>% anti_join(allTotalPAS,by = "pasNum")

ggplot(NuclearSpec,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Nuclear  Specific PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

Total Specific:
```{r}
TotalSpec=allTotalPAS %>% anti_join(allNuclearPAS,by = "pasNum")

ggplot(TotalSpec,aes(x=loc,fill=withAnno)) + geom_histogram(stat="count") + labs(title = "Total  Specific PAS by annotated PAS within 10bp") +scale_fill_brewer(palette = "Dark2")
```

