---
title: "Conservation of PAS"
author: "Briana Mittleman"
date: "1/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I want to study the conservation of the PAS. I will compare PAS with QTLs and those without. I will use PhyloP score. PhyloP scores for 100 vertibrates are available on the genome browser.  

```{r}
library(workflowr)
library(tidyverse)
library(ggpubr)
```

```{bash,eval=F}
mkdir ../data/phylop/
```

PhyloP: Column #1 contains a one-based position coordinate. Column #2 contains a score showing the posterior probability that the phylogenetic hidden Markov model (HMM) of phastCons is in its most conserved state at that base position. I will use pybigwig to extract the regions I care about.  

I will look at the 200 basepairs around each PAS.  

```{r}
pas=read.table("../data/PAS/APAPAS_GeneLocAnno.5perc.sort.bed",col.names = c("chr","start","end","name","score","strand")) %>% mutate(newStart=end-100,newEnd=end+100) %>% select(chr,newStart,newEnd, name)

write.table(pas,"../data/phylop/PAS_regions.txt",col.names = F,row.names = F,quote = F,sep="\t")
```


```{bash,eval=F}
python extractPACmeanPhyloP.py
```

Add information about qtl or not:  

```{r}
nucQTL=read.table("../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt",header = T) 
```

```{r}
phylores=read.table("../data/phylop/PAS_phyloP.txt", col.names = c("chr","start","end","name", "phyloP"), stringsAsFactors = F) %>% drop_na() %>% separate(name,into=c("pasnum","geneid"), sep=":") %>% mutate(PAS=paste("peak",pasnum,sep="")) %>% mutate(HasQTL=ifelse(PAS %in% nucQTL$Peak, "Yes","No"))

```
41,810 - 41649

lost 161 to NAs

Plot:  

```{r}
ggplot(phylores,aes(x=phyloP, by=HasQTL, fill=HasQTL)) +geom_density(alpha=.5)

ggplot(phylores,aes(y=phyloP, x=HasQTL,fill=HasQTL)) + geom_boxplot() + stat_compare_means()+ scale_fill_brewer(palette = "Dark2", name="Signficant")
```

With a QTL is lower score.  Look at enrichment for negative  

```{r}
x=nrow(phylores %>% filter(HasQTL=="Yes", phyloP<0))
m= nrow(phylores %>% filter(phyloP<0))
n=nrow(phylores %>% filter(phyloP>=0))
k=nrow(phylores %>% filter(HasQTL=="Yes"))


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))

#actual:
x

#pval
phyper(x,m,n,k,lower.tail=F)
```

Enriched for regions expected to be rapidly evolving.  



Try with smaller regions:  

```{r}
pas100=read.table("../data/PAS/APAPAS_GeneLocAnno.5perc.sort.bed",col.names = c("chr","start","end","name","score","strand")) %>% mutate(newStart=end-50,newEnd=end+50) %>% select(chr,newStart,newEnd, name)

write.table(pas,"../data/phylop/PAS100_regions.txt",col.names = F,row.names = F,quote = F,sep="\t")
```
```{bash,eval=F}
python extactPAS100meanphyloP.py
```

```{r}
phylores100=read.table("../data/phylop/PAS100_phyloP.txt", col.names = c("chr","start","end","name", "phyloP"), stringsAsFactors = F) %>% drop_na() %>% separate(name,into=c("pasnum","geneid"), sep=":") %>% mutate(PAS=paste("peak",pasnum,sep="")) %>% mutate(HasQTL=ifelse(PAS %in% nucQTL$Peak, "Yes","No"))

```


```{r}
ggplot(phylores100,aes(x=phyloP, by=HasQTL, fill=HasQTL)) +geom_density(alpha=.5)

ggplot(phylores100,aes(y=phyloP, x=HasQTL,fill=HasQTL)) + geom_boxplot() + stat_compare_means()+ scale_fill_brewer(palette = "Dark2", name="Signficant")
```

```{r}
x=nrow(phylores100 %>% filter(HasQTL=="Yes", phyloP<0))
m= nrow(phylores100 %>% filter(phyloP<0))
n=nrow(phylores100 %>% filter(phyloP>=0))
k=nrow(phylores100 %>% filter(HasQTL=="Yes"))


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))

#actual:
x

#pval
phyper(x,m,n,k,lower.tail=F)
```
Exact same results.  
