---
title: "eQTL effect size for Nuclear specific APAqtls"
author: "Briana Mittleman"
date: "6/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(workflowr)
```

I am interested in seeing if the nuclear specific apaQTL snps have higher eQTL effect sizes than the shared apaQTLs. This comes from my expectation that the nuclear specific apaQTLs are more likely to be eQTLs. 

I am using the nuclear specific apa QTLs including those that are not tested in total from [this analysis](NuclearSpecIncludeNotTested.Rmd).

```{bash,eval=F}
python phenoQTLfromlist.py ../data/NucSpeceQTLeffect/NucSpecApaQTL_nottestinc.txt RNA ../data/NucSpeceQTLeffect/NuclearSpecficApaQTL_eQTLRes.txt
python phenoQTLfromlist.py ../data/NucSpeceQTLeffect/SharedApaQTL_nottestinc.txt RNA ../data/NucSpeceQTLeffect/SharedApaQTL_eQTLRes.txt

```


Evaluate results:
```{r}
eqtl4nucqtl=read.table("../data/NucSpeceQTLeffect/NuclearSpecficApaQTL_eQTLRes.txt", col.names = c("gene", "snp", "dist", "pval", "effect"),stringsAsFactors = F) %>% mutate(set="Nuclear")

eqtl4sharedqtl=read.table("../data/NucSpeceQTLeffect/SharedApaQTL_eQTLRes.txt", col.names = c("gene", "snp", "dist", "pval", "effect"),stringsAsFactors = F)%>% mutate(set="Shared")


eQTLsboth=eqtl4nucqtl %>% bind_rows(eqtl4sharedqtl)
```



Plot:

```{r}
ggplot(eQTLsboth,aes(x=set, y=abs(effect)))+ geom_boxplot()
ggplot(eQTLsboth,aes(x=set, y=pval))+ geom_boxplot()+ geom_jitter() 
```
No effect. What if we jsut look at nominally sig snps.  

```{r}
eQTLsbothSig=eQTLsboth %>% filter(pval<.05)

ggplot(eQTLsbothSig,aes(x=set, y=abs(effect),fill=set))+ geom_boxplot() + geom_jitter() 
```

```{r}

t.test(abs(effect) ~ set, data = eQTLsbothSig, alternative="greater")
```



```{bash,eval=F}
python phenoQTLfromlist.py ../data/NucSpeceQTLeffect/NucSpecApaQTL_nottestinc.txt Protein ../data/NucSpeceQTLeffect/NucSpecApaQTL_nottestinc_pQTLRes.txt
python phenoQTLfromlist.py ../data/NucSpeceQTLeffect/SharedApaQTL_nottestinc.txt Protein ../data/NucSpeceQTLeffect/SharedApaQTLnottestedinc_pQTLRes.txt
```

```{r}
pqtl4nucqtl=read.table("../data/NucSpeceQTLeffect/NucSpecApaQTL_nottestinc_pQTLRes.txt", col.names = c("gene", "snp", "dist", "pval", "effect"),stringsAsFactors = F) %>% mutate(set="Nuclear")

pqtl4sharedqtl=read.table("../data/NucSpeceQTLeffect/SharedApaQTLnottestedinc_pQTLRes.txt", col.names = c("gene", "snp", "dist", "pval", "effect"),stringsAsFactors = F)%>% mutate(set="Shared")


pQTLsboth=pqtl4nucqtl %>% bind_rows(pqtl4sharedqtl)
```


```{r}
ggplot(pQTLsboth,aes(x=set, y=abs(effect)))+ geom_boxplot()
ggplot(pQTLsboth,aes(x=set, y=pval))+ geom_boxplot()+ geom_jitter() 
```

```{r}
pQTLsbothSig=pQTLsboth %>% filter(pval<.05)

ggplot(pQTLsbothSig,aes(x=set, y=abs(effect),fill=set))+ geom_boxplot() + geom_jitter() 
```
Not significant.  
