---
title: "Explained and unexplained eQTLs at gene level"
author: "Briana Mittleman"
date: "6/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

I want to ask if the unexplained and explaiend eQTLs are in the same or different genes.  

```{r}
explainedGenes=read.table("../data/Li_eQTLs/explained_FDR10.sort_FIXED.txt", stringsAsFactors = F, col.names = c("chr", "snp", "gene") )%>% select(gene) %>% unique() 

write.table(explainedGenes, file="../data/Li_eQTLs/explainedEgenes.txt", col.names = F, row.names = F, quote = F, sep="\t")

UNexplainedGenes=read.table("../data/Li_eQTLs/unexplained_FDR10.sort_FIXED.txt", stringsAsFactors = F, col.names = c("chr", "snp", "gene") )%>% select(gene) %>% unique()
write.table(UNexplainedGenes, file="../data/Li_eQTLs/UnexplainedEgenes.txt",col.names = F, row.names = F, quote = F, sep="\t")
```


Make these vectors:

```{r}
explainedGenesVec=explainedGenes$gene
UNexplainedGenesVec=UNexplainedGenes$gene



```


Overlap:

```{r}

both <- UNexplainedGenesVec[UNexplainedGenesVec %in% explainedGenesVec]
both
```


Plot the permuted pvalues for apa in 3 seperate lines. not eGenes, unexplained egenes, explained egenes 


I want to make a script that takes these genes and a fraction and will give only the permuted apa values for that set. I will also make a script that will return the values for genes in neither set.  


```{bash,eval=F}
mkdir ../data/ApaByEgene
python subsetpermAPAwithGenelist.py ../data/Li_eQTLs/explainedEgenes.txt Total ../data/ApaByEgene/TotalApaexplainedeGenes.txt

python subsetpermAPAwithGenelist.py ../data/Li_eQTLs/UnexplainedEgenes.txt Total ../data/ApaByEgene/TotalApaUnexplainedeGenes.txt

python subsetpermAPAwithGenelist.py ../data/Li_eQTLs/explainedEgenes.txt Nuclear ../data/ApaByEgene/NuclearApaexplainedeGenes.txt

python subsetpermAPAwithGenelist.py ../data/Li_eQTLs/UnexplainedEgenes.txt Nuclear ../data/ApaByEgene/NuclearApaUnexplainedeGenes.txt


python subsetApanoteGene.py Total ../data/ApaByEgene/TotalApaNOTeGene.txt
python subsetApanoteGene.py Nuclear ../data/ApaByEgene/NuclearApaNOTeGene.txt

```



Make QQplots with these:  

###Total
```{r}
tot.notE=read.table("../data/ApaByEgene/TotalApaNOTeGene.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
tot.ex=read.table("../data/ApaByEgene/TotalApaexplainedeGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
tot.un=read.table("../data/ApaByEgene/TotalApaUnexplainedeGenes.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval") )
tot.un=na.omit(tot.un)
```


```{r}
qqplot(-log10(runif(nrow(tot.notE))), -log10(tot.notE$bpval), xlab="-log10(Uniform)", ylab="-log10(beta pval)", main="Total Apa")
points(sort(-log10(runif(nrow(tot.ex)))), sort(-log10(tot.ex$bpval)),col= alpha("Red"))
points(sort(-log10(runif(nrow(tot.un)))), sort(-log10(tot.un$bpval)),col= alpha("Blue"))
abline(0,1)

legend("topleft", legend=c("Not eGenes", "Explained eGenes", "Unexplained eGenes"),col=c("black", "red", "blue"), pch=16,bty = 'n')

```
```{r}
wilcox.test(tot.notE$bpval,tot.ex$bpval,alternative = "greater")
wilcox.test(tot.notE$bpval,tot.un$bpval, alternative = "greater")
wilcox.test(tot.un$bpval,tot.ex$bpval, alternative = "less")
```



###Nuclear  

```{r}
nuc.notE=read.table("../data/ApaByEgene/NuclearApaNOTeGene.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
nuc.ex=read.table("../data/ApaByEgene/NuclearApaexplainedeGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
nuc.un=read.table("../data/ApaByEgene/NuclearApaUnexplainedeGenes.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval") )
nuc.un=na.omit(nuc.un)
```


```{r}
qqplot(-log10(runif(nrow(nuc.notE))), -log10(nuc.notE$bpval), xlab="-log10(Uniform)", ylab="-log10(beta pval)", main="Nuclear Apa")
points(sort(-log10(runif(nrow(nuc.ex)))), sort(-log10(nuc.ex$bpval)),col= alpha("Red"))
points(sort(-log10(runif(nrow(nuc.un)))), sort(-log10(nuc.un$bpval)),col= alpha("Blue"))
abline(0,1)
legend("topleft", legend=c("Not eGenes", "Explained eGenes", "Unexplained eGenes"),col=c("black", "red", "blue"), pch=16,bty = 'n')

```

```{r}
wilcox.test(nuc.notE$bpval,nuc.ex$bpval,alternative="greater")
wilcox.test(nuc.notE$bpval,nuc.un$bpval,alternative="greater")
wilcox.test(nuc.un$bpval,nuc.ex$bpval,alternative="less")
```


