---
title: "Most credible set PAS"
author: "Briana Mittleman"
date: "6/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(workflowr)
```

I want to get the most credible intronic PAS set I can. To do this I can use analysis I have previously done. These are those that have a signal site and have more RNA seq reads upstream. This mean I am looking at those in the total fraction.  

1. [Signal site analyis](signalsiteanalysis.html)  

2. [RNA evidence upstream](intonRNAratio.html)  



```{r}
signalPAS=read.table("../data/PAS/PASwSignalSite.txt", header  =T, stringsAsFactors = F) %>% filter(loc=="intron")

```


```{r}
RNAupstream=read.table(file="../data/intronRNAratio/TotalPAS_MoreUpstreamRNAreads.txt", header = T, stringsAsFactors = F) 
```


Join these together on the PAS gene and loc

```{r}
PAS_signalandRNA=RNAupstream %>% inner_join(signalPAS, by=c("PAS", "gene", "loc"))


```


I can look at these in IGV.  


Are any of these total QTLs?

```{r}
totQTL=read.table("../data/apaQTLs/Total_apaQTLs_5fdr.txt", header = T, stringsAsFactors = F) %>% dplyr::rename("PAS"=Peak)
```


Filter join the PAS set with the QTLs 

```{r}
highcredwQTL=PAS_signalandRNA %>%  semi_join(totQTL, by= "PAS")
```

