---
title: "Locus zoom plots"
author: "Briana Mittleman"
date: "2/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

The locuszoom package is in:  
/scratch/midway2/brimittleman/locuszoom


Run example in /example. Use /scratch/midway2/brimittleman/locuszoom/bin/locuszoom --metal Kathiresan_2009_HDL.txt --refgene FADS1 --pop AFR --build hg19 --source 1000G_March2012

Modules to load:
```{bash,eval=F}
module load R
module load plink
module load htslib
```

File formats needed:  

MarkerName	P-value	Weight

This will be the snp and the nominal pvalues.   

--refsnp <your snp> --flank 500kb  --metal 

 title="My region" geneFontSize=1.1
 
 I can do one then loop it.  
 
 
```{bash,eval=F}
mkdir ../data/LZ_both/ 
mkdir ../data/LZ_both/NuclearAssoc
mkdir ../data/LZ_both/RNAAssoc

```
 
 /scratch/midway2/brimittleman/locuszoom/
 
 I need a file with all of the apaQTL associations for the PAS of interest. Instead of genes of interest I can work with PAS of interest.  
 
```{r}
NuclearQTLs=read.table("../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt", head=T) %>% select(Peak) %>% unique()

write.table(NuclearQTLs,"../data/apaQTLs/NuclearQTL_PAS.txt", col.names = F, row.names = F, quote = F)
```
 
Select these from  /project2/gilad/briana/apaQTL/data/apaQTLNominal_4pc/APApeak_Phenotype_GeneLocAnno.Total.5perc.fc.gz.qqnorm_AllChrom.txt  

I will have to do this seperatly in case there are snp overlaps.  
 
Make the file rsid and pval for each. 

Write a python script that will take in the PAS and write out the file. I need to do this by gene. I will do the genes with the apaQTL genes. 

```{r}
NucQ=read.table("../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt", head=T) %>% select(Peak, sid, Gene)  

write.table(NucQ,"../data/apaQTLs/NuclearQTLs_PeakSNP.txt", quote=F, row.names = F, col.names = F)
```


```{bash,eval=F}
 sbatch ExtractPASforLZ_run.sh 
 #rna file
 # ../data/molQTLs/fastqtl_qqnorm_RNAseq_phase2.fixed.nominal.AllNomRes.GeneName.txt
 #filter this
 sbatch extractGeneLZfiles.sh
 
 
```


Test one with RNA.  rs7076314

Do this in ../data/LZ_both/ 
/scratch/midway2/brimittleman/locuszoom/bin/locuszoom --metal  RNA_Res.txt --refsnp rs7076314 --flank 100kb title="rs7076314" geneFontSize=1.1 --pop AFR --build hg19 --source 1000G_March2012


I need a file with the snp and PAS.  

I can plot nuclear PAS  wit hthe --bed-tracks <your bed file>  

```{r}
NuclearPAS=read.table("../data/PAS/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.sort.bed",col.names = c("chr","start", "end", "name", "score", "strand")) %>% mutate(Col="PAS", score2=".") %>% select(chr, start,end, Col, score2, strand)

write.table(NuclearPAS, "../data/LZ_both/NuclearPAS.bed", col.names = F, row.names = F, quote=F, sep="\t")
```

If I want to add GWAS, I can add --gwas-cat whole-cat_significant-only  


Test 

```{bash,eval=F}
/scratch/midway2/brimittleman/locuszoom/bin/locuszoom --metal  RNAAssoc/ABTB1RNA_Res.txt --refsnp rs60645978 --flank 100kb title="rs60645978" geneFontSize=1.1 --pop AFR --build hg19 --source 1000G_March2012 --bed-tracks NuclearPAS.bed  theme="publication"

#peak91979
/scratch/midway2/brimittleman/locuszoom/bin/locuszoom --metal  NuclearAssoc/peak91979_NuclearResults4LZ.txt --refsnp rs60645978 --flank 100kb title="rs60645978" geneFontSize=1.1 --pop AFR --build hg19 --source 1000G_March2012 --bed-tracks NuclearPAS.bed  theme="publication"

```

Wrap all 
run in the dir i want the data to come out
```{bash,eval=F}
cd ../data/LZ_both/RNAAssoc
sbatch CreateRNALZnucAPAqtls.sh

cd ../data/LZ_both/NuclearAssoc
sbatch CreateAPALZnucAPAqtls.sh
```

All of the plots moved to 

/project2/gilad/briana/apaQTL/data/LZ_both/Both_Plots


They are named by the region.  

I can look at these.  


```{r}
QTLs=read.table("../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt", head=T) 
```

By eye 41 colocalize: 

rs74064859
rs7523082
rs72786277
rs12468045
rs7587301
rs1859925
rs2067987
rs4974091
rs142895723
rs6785658
rs3181141
rs13327891
rs62291862
rs114653275
rs568902
rs2287371
rs113268660
rs73527330
rs145898840
rs12339205
rs73251478
rs7071284
rs17112791
rs7897454
rs41291460
rs41291462
rs3740400
rs7904973 (c10orf88) 
rs10840038
rs10891659
rs6650507
rs1956367
rs448113
rs10149617
rs4902346
rs75664871
rs113727694
rs115018779
rs78139339
rs6085202
rs2184159  

of 524 succesful  

(41/524)
