---
title: "Mechanism for each expression independent apaQTL"
author: "Briana Mittleman"
date: "2/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I will pull in the apaQTLs that are pQTLs or riboQTLs but are not eQTLs. I will look at their UTRs to think about mechanisms.  

```{r}
library(workflowr)
library(tidyverse)
```

```{r}
QTLs=read.table("../data/ExpressionIndependentapaQTLs.txt",header = T,stringsAsFactors = F)



QTLs %>% select(Gene, PAS_ID, apa_EffectSize,protein_EffectSize,protein_pvalue,ribo_EffectSize,ribo_pvalue)
```

DIP2B: QTL in proximal UTR, hsa-miR-143-3p, hsa-miR-6088, hsa-miR-4770, miRNAs are upstream of first 

The K box (cTGTGATa) is present in one or more copies in many of these 3'UTRs and mediates negative post-transcriptional regulation, mainly effected by decreased transcript levels.

C12orf73: 2nd of 3 UTR PAS:
- POLR2A binding sites 
-hsa-miR-9-5p 

OAS1:  

IFI44L: between 2 UTR FOS binding site 


STXBP3: PUM1 binding in 3' UTR, translational regulator of specific mRNAs by binding to their 3' untranslated regions, reported kbox, many miRNAs conserved (miR-15-5p/16-5p/195-5p/424-5p/497-5p)  

RBBP5: extended isoform,reported Brd-Box (Brd)  The Brd box (AGCTTTA) is present in one or more copies in many of these 3'UTRs and mediates negative post-transcriptional regulation by affecting transcript stability and translational efficiency. hsa-miR-338-3p

PSMF1: UNR binding site: In humans, UNR has been implicated in the destabilization of c-fos mRNA (Chang et al., 2004) and the activation of translation driven by the IRESs of a number of transcripts, including c-myc, miR-103-3p/107

MGME1- RBP binding for AQR, NIPBL, FUS (no DB results)

FAM207A- (no DB results)

APOL2- reported UNR binding site, SAFB binding between PAS in UTR

MPHOSPH10-

SACM1L-miR-181-5p

EIF2A- annotated Brd-Box (Brd) site between middle 2 utr binding sites, reported kbox as well 


PCYT1A- (no DB results) many binding sites between annotated utr and end PAS, miR-142-3p.2

HTT: BRD-BOX, hsa-miR-216b-5p


STAP1- LTR/ERVL-MaLR repete element 


NUP155-KBOX, BRD BOX , GLUT1  element, miR-199-3p, hsa-miR-3129-5p

ERAP1-MBE element - related to temporal regulation ,hsa-miR-4770, hsa-miR-143-3p, hsa-miR-6088

SQSTM1- FLAM_A - SINE/Alu repeate, GY-BOX- involved in RNA duplexes, hsa-miR-7-5p

BTBD9-  GY-BOX, conserved hsa-miR-212-5p binding site , hsa-miR-212-5p

STX17- GY-BOX, BRD BOX, k-box, miR-15-5p/16-5p/195-5p/424-5p/497-5p

Lit search for BRD box and K box only has results for human, info about regulation in drosophila 


