---
title: "Compare new and old peaks"
author: "Briana Mittleman"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(workflowr)
```

These results have 30k more PAS than the previous runs. I also see a confusing shift in mean usage for all of the PAS. I want to compare the distribution of usage for different sets of individuals to see if there is something inherently different about the 15 new individuals.  


##New vs old peaks

I want to compare the usage of the new peaks compared to the overall mean usage. To do this I need to seperate the new and old PAS. 

```{r}
newPAS5perc=read.table("../data/PAS/APAPAS_GeneLocAnno.5perc.bed", stringsAsFactors = F, col.names = c("chr", "start","end", "ID", "score", "strand"))
oldPAS5perc=read.table("../../threeprimeseq/data/peaks4DT/APAPAS_5percCov_fixedStrand.bed", stringsAsFactors = F, col.names = c("chr", "start", "end", "ID", "score", "strand"))

uniqnew=newPAS5perc %>% semi_join(oldPAS5perc, by=c("chr", "start", "end"))
```

Pull in the usage of the peaks:  

###Total
```{r}
totalPeakUs=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.fc", header = T, stringsAsFactors = F) %>% separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% separate(id, sep="_", into=c("gene", "loc", "strand", "peak"))


ind=colnames(totalPeakUs)[8:dim(totalPeakUs)[2]]
totalPeakUs_CountNum=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.CountsOnlyNumeric", col.names = ind)


#numeric with anno
totalPeak=as.data.frame(cbind(totalPeakUs[,1:7], totalPeakUs_CountNum))

totalPeakUs_CountNum_mean=rowMeans(totalPeakUs_CountNum)

#append mean to anno
TotalPeakUSMean=as.data.frame(cbind(totalPeakUs[,1:7],mean=totalPeakUs_CountNum_mean))

```

```{r}
uniqnewPasnum=uniqnew  %>% separate(ID ,into=c("peaknum", "geneloc"),sep=":") %>% mutate(peak=paste("peak", peaknum, sep="")) %>% select(peak)
```

Filter these inthe mean usage: 

```{r}
TotalPeakUSMeanClass= TotalPeakUSMean %>% mutate(New=ifelse(peak %in% uniqnewPasnum$peak,"new", "original")) %>% mutate(Cutoff=ifelse(mean>=.05, "Yes", "No"))

mean(TotalPeakUSMean$mean)
```


Plot:

```{r}
ggplot(TotalPeakUSMeanClass, aes(y=mean,x=New)) + geom_violin() + geom_hline(yintercept = mean(TotalPeakUSMean$mean), col="red") + facet_grid(~Cutoff)

```

This shows me the new peaks are not the peaks that barely passed the cuttoff before. These peaks cover the distribution of usage.  

write out file with information about new and old peaks  
```{r}
Peak_newOld=TotalPeakUSMeanClass %>% select(-mean)
write.table(Peak_newOld, file="../data/peaks_5perc/NewVOldPeaks.txt", col.names = T, row.names = F, quote=F)
```


##Distirubution in different sets of 15 ind  

I want to see if the new 15 individuals are driving the change in the peak mean distribution. I want to make a function that take a vector of individuals, filters the usage dataframe and plots the histogram.  


###Total: 
First I will upload the usage dataframe. 

```{r}
totCounts=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.fc", stringsAsFactors = F, header = T)
ind=colnames(totCounts)[2:55]
totUsage=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.CountsOnlyNumeric", stringsAsFactors = F, header = F,col.names = ind)



batch1.2.3=read.table("../data/MetaDataSequencing.txt", header=T,stringsAsFactors = F)%>% filter(fraction=="total") %>%  select(line, batch) %>% filter(batch != 4)
oldind=batch1.2.3$line

batch4=read.table("../data/MetaDataSequencing.txt", header=T,stringsAsFactors = F)%>% filter(fraction=="total") %>%  select(line, batch) %>% filter(batch == 4)

newInd=batch4$line

  

```



```{r}
UsageHist= function(indVec,title,totUsage=totUsage){
  totUsage_ind=totUsage %>% select(indVec)
  meanVec=rowMeans(totUsage_ind)
  hist(meanVec, main=title,xlab="Mean Usage")
}

```


RUn this for different itterations of individuals:


```{r}

par(mfrow=c(1,2))
#old ind
UsageHist(indVec=newInd,title="Total Usage New",totUsage=totUsage)
UsageHist(indVec=oldind,title="Total Usage Old",totUsage=totUsage)
```
Pick 15 random individuals from old: 


```{r}

sampl1=sample(oldind, 15)
sampl2=sample(oldind, 15)
sampl3=sample(oldind, 15)
sampl4=sample(oldind, 15)


par(mfrow=c(2,2))
UsageHist(indVec=sampl1,title="Total Usage Sample 15 Old",totUsage=totUsage)
UsageHist(indVec=sampl2,title="Total Usage Sample 15 Old",totUsage=totUsage)
UsageHist(indVec=sampl3,title="Total Usage Sample 15 Old",totUsage=totUsage)
UsageHist(indVec=sampl4,title="Total Usage Sample 15 Old",totUsage=totUsage)
```





###Nuclear  

```{r}
nucCounts=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Nuclear.fc", stringsAsFactors = F, header = T)
ind=colnames(nucCounts)[2:55]
nucUsage=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Nuclear.CountsOnlyNumeric", stringsAsFactors = F, header = F,col.names = ind)

```



```{r}
UsageHist_nuc= function(indVec,title,nucUsage=nucUsage){
  nucUsage_ind=nucUsage %>% select(indVec)
  meanVec=rowMeans(nucUsage_ind)
  hist(meanVec, main=title,xlab="Mean Usage")
}

```


```{r}

par(mfrow=c(1,2))
UsageHist_nuc(indVec=newInd,title="Nuclear Usage New",nucUsage=nucUsage)
UsageHist_nuc(indVec=oldind,title="Nuclear Usage Old",nucUsage=nucUsage)
```

```{r}

par(mfrow=c(2,2))
UsageHist_nuc(indVec=sampl1,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl2,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl3,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl4,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
```
```

