---
title: "Compare new and old peaks"
author: "Briana Mittleman"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(workflowr)
```

These results have 30k more PAS than the previous runs. I also see a confusing shift in mean usage for all of the PAS. I want to compare the distribution of usage for different sets of individuals to see if there is something inherently different about the 15 new individuals.  


##New vs old peaks

I want to compare the usage of the new peaks compared to the overall mean usage. To do this I need to seperate the new and old PAS. 

```{r}
newPAS5perc=read.table("../data/PAS/APAPAS_GeneLocAnno.5perc.bed", stringsAsFactors = F, col.names = c("chr", "start","end", "ID", "score", "strand"))
oldPAS5perc=read.table("../../threeprimeseq/data/peaks4DT/APAPAS_5percCov_fixedStrand.bed", stringsAsFactors = F, col.names = c("chr", "start", "end", "ID", "score", "strand"))

uniqnew=newPAS5perc %>% anti_join(oldPAS5perc, by=c("chr", "start", "end"))
```

Pull in the usage of the peaks:  

###Total
```{r}
totalPeakUs=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.fc", header = T, stringsAsFactors = F) %>% separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% separate(id, sep="_", into=c("gene", "loc", "strand", "peak"))


ind=colnames(totalPeakUs)[8:dim(totalPeakUs)[2]]
totalPeakUs_CountNum=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.CountsOnlyNumeric", col.names = ind)


#numeric with anno
totalPeak=as.data.frame(cbind(totalPeakUs[,1:7], totalPeakUs_CountNum))

totalPeakUs_CountNum_mean=rowMeans(totalPeakUs_CountNum)

#append mean to anno
TotalPeakUSMean=as.data.frame(cbind(totalPeakUs[,1:7],mean=totalPeakUs_CountNum_mean))

```

```{r}
uniqnewPasnum=uniqnew  %>% separate(ID ,into=c("peaknum", "geneloc"),sep=":") %>% mutate(peak=paste("peak", peaknum, sep="")) %>% select(peak)
```

Filter these inthe mean usage: 

```{r}
TotalPeakUSMeanClass= TotalPeakUSMean %>% mutate(New=ifelse(peak %in% uniqnewPasnum$peak,"new", "original")) %>% mutate(Cutoff=ifelse(mean>=.05, "Yes", "No"))

mean(TotalPeakUSMean$mean)
```


Plot:

```{r}
ggplot(TotalPeakUSMeanClass, aes(y=mean,x=New)) + geom_violin() + geom_hline(yintercept = mean(TotalPeakUSMean$mean), col="red") +  geom_hline(yintercept = .05, col="Blue")

```


```{r}
TotalPeakUSMeanClass_newonly= TotalPeakUSMeanClass %>% filter(New=="new")

ggplot(TotalPeakUSMeanClass_newonly, aes(y=mean, x="")) + geom_violin() + geom_hline(yintercept = mean(TotalPeakUSMean$mean), col="red") +  geom_hline(yintercept = .05, col="Blue") + labs(x="", y="Mean Usage", title="Mean Usage of New PAS")


```
```{r}
ggplot(TotalPeakUSMeanClass_newonly, aes(x=mean)) + geom_histogram(bins=50) + geom_vline(xintercept = mean(TotalPeakUSMean$mean), col="red")
```

This shows me the new peaks are the peaks that barely passed the cuttoff before.  

write out file with information about new and old peaks  
```{r}
Peak_newOld=TotalPeakUSMeanClass %>% select(-mean)
write.table(Peak_newOld, file="../data/peaks_5perc/NewVOldPeaks.txt", col.names = T, row.names = F, quote=F)
```


##Distirubution in different sets of 15 ind  

I want to see if the new 15 individuals are driving the change in the peak mean distribution. I want to make a function that take a vector of individuals, filters the usage dataframe and plots the histogram.  


###Total: 
First I will upload the usage dataframe. 

```{r}
totCounts=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.fc", stringsAsFactors = F, header = T)
ind=colnames(totCounts)[2:55]
totUsage=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Total.CountsOnlyNumeric", stringsAsFactors = F, header = F,col.names = ind)



batch1.2.3=read.table("../data/MetaDataSequencing.txt", header=T,stringsAsFactors = F)%>% filter(fraction=="total") %>%  select(line, batch) %>% filter(batch != 4)
oldind=batch1.2.3$line

batch4=read.table("../data/MetaDataSequencing.txt", header=T,stringsAsFactors = F)%>% filter(fraction=="total") %>%  select(line, batch) %>% filter(batch == 4)

newInd=batch4$line

  

```



```{r}
UsageHist= function(indVec,title,totUsage=totUsage){
  totUsage_ind=totUsage %>% select(indVec)
  meanVec=rowMeans(totUsage_ind)
  hist(meanVec, main=title,xlab="Mean Usage")
}

```


RUn this for different itterations of individuals:


Pick 15 random individuals from old: 


```{r}

sampl1=sample(oldind, 15)
sampl2=sample(oldind, 15)
sampl3=sample(oldind, 15)
sampl4=sample(oldind, 15)

png("../output/newtot.png")
par(mfrow=c(3,2))
UsageHist(indVec=newInd,title="Total Usage New (15ind)",totUsage=totUsage)
UsageHist(indVec=oldind,title="Total Usage Old (39 ind)",totUsage=totUsage)
UsageHist(indVec=sampl1,title="Total Usage Sample 15 Old",totUsage=totUsage)
UsageHist(indVec=sampl2,title="Total Usage Sample 15 Old",totUsage=totUsage)
UsageHist(indVec=sampl3,title="Total Usage Sample 15 Old",totUsage=totUsage)
UsageHist(indVec=sampl4,title="Total Usage Sample 15 Old",totUsage=totUsage)
dev.off()
```





###Nuclear  

```{r}
nucCounts=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Nuclear.fc", stringsAsFactors = F, header = T)
ind=colnames(nucCounts)[2:55]
nucUsage=read.table("../data/phenotype/APApeak_Phenotype_GeneLocAnno.Nuclear.CountsOnlyNumeric", stringsAsFactors = F, header = F,col.names = ind)

```



```{r}
UsageHist_nuc= function(indVec,title,nucUsage=nucUsage){
  nucUsage_ind=nucUsage %>% select(indVec)
  meanVec=rowMeans(nucUsage_ind)
  hist(meanVec, main=title,xlab="Mean Usage")
}

```



```{r}
png("../output/newnuc.png")
par(mfrow=c(3,2))
UsageHist_nuc(indVec=newInd,title="Nuclear Usage New (15ind)",nucUsage=nucUsage)
UsageHist_nuc(indVec=oldind,title="Nuclear Usage Old (39ind)",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl1,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl2,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl3,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
UsageHist_nuc(indVec=sampl4,title="Nuclear Usage Sample 15 Old",nucUsage=nucUsage)
dev.off()
```


Old 

```{r}

oldtotalCount=read.table("../../threeprimeseq/data/phenotypes_filtPeakTranscript_noMP_GeneLocAnno/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno.fc",header=T,stringsAsFactors = F)
indOld=colnames(oldtotalCount)[2:56]
oldtotalUsage=read.table("../../threeprimeseq/data/phenotypes_filtPeakTranscript_noMP_GeneLocAnno/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno.CountsOnlyNumeric.txt",col.names = indOld,stringsAsFactors = F)


```


```{r}
png("../output/oldtot.png")
par(mfrow=c(3,2))
UsageHist(indVec=newInd,title="Old total Usage (15ind)",totUsage=oldtotalUsage)
UsageHist(indVec=oldind,title="Old total Usage (39ind)",totUsage=oldtotalUsage)
UsageHist(indVec=sampl1,title="Old total Usage sample 15 ind",totUsage=oldtotalUsage)
UsageHist(indVec=sampl2,title="Old total Usage sample 15 ind",totUsage=oldtotalUsage)
UsageHist(indVec=sampl3,title="Old total Usage sample 15 ind",totUsage=oldtotalUsage)
UsageHist(indVec=sampl4,title="Old total Usage sample 15 ind",totUsage=oldtotalUsage)
dev.off()
```




```{r}

oldnuclearCount=read.table("../../threeprimeseq/data/phenotypes_filtPeakTranscript_noMP_GeneLocAnno/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno.fc",header=T,stringsAsFactors = F)
indOldN=colnames(oldnuclearCount)[2:56]
oldnuclearUsage=read.table("../../threeprimeseq/data/phenotypes_filtPeakTranscript_noMP_GeneLocAnno/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno.CountsOnlyNumeric.txt",col.names = indOldN,stringsAsFactors = F)

```

```{r}
png("../output/oldnuc.png")
par(mfrow=c(3,2))
UsageHist_nuc(indVec=newInd,title="Old nuclear Usage (15ind)",nucUsage=oldnuclearUsage)
UsageHist_nuc(indVec=oldind,title="Old nuclear Usage (39ind)",nucUsage=oldnuclearUsage)
UsageHist_nuc(indVec=sampl1,title="Old nuclear Usage sample 15 ind",nucUsage=oldnuclearUsage)
UsageHist_nuc(indVec=sampl2,title="Old nuclear Usage sample 15 ind",nucUsage=oldnuclearUsage)
UsageHist_nuc(indVec=sampl3,title="Old nuclear Usage sample 15 ind",nucUsage=oldnuclearUsage)
UsageHist_nuc(indVec=sampl4,title="Old nuclear Usage sample 15 ind",nucUsage=oldnuclearUsage)
dev.off()
```



##Usage of new peaks in old data  
I have the old usage. I want to filter the new peaks from this:  
```{r}

oldtotusage_anno=as.data.frame(cbind(chrom=oldtotalCount$chrom,oldtotalUsage )) %>% separate(chrom, into=c("chr", "start", "end", "peakID"), sep=":")

TotalPeakUSMeanClass_newonly=TotalPeakUSMeanClass %>% filter(New=="new")

oldtotusage_anno_new=oldtotusage_anno %>% semi_join(TotalPeakUSMeanClass_newonly, by=c("chr", "start", "end")) %>% select(-chr, -start,-end,-peakID)


oldtotusage_anno_new_mean=rowMeans(oldtotusage_anno_new)

plot(oldtotusage_anno_new_mean, main="Total Usage of new peaks in old data", ylab="old usage means Percent")

length(oldtotusage_anno_new_mean)
```

This shows me that we called 14 thousand of the new peaks in the old set but they were all super low coverage.  



Where are the new peaks, genes with a lot of peaks? or genes with less peaks  



```{r}
newPAS5perc_pergene=newPAS5perc %>% separate(ID, into=c("peaknum", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc"),sep="_") %>% group_by(gene) %>% summarise(nPeak=n())

ggplot(newPAS5perc_pergene,aes(x=nPeak)) + geom_histogram(bins=100)


nrow(newPAS5perc_pergene)
```
Look at which genes the new peaks are in.

```{r}
uniqnew_genes=uniqnew %>% separate(ID, into=c("peaknum", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "loc"),sep="_") %>% group_by(gene) %>% summarise(npeakadded=n())


ggplot(uniqnew_genes, aes(x=npeakadded)) + geom_histogram(bins=100) + labs(title="Number of peaks added per gene\n(added peak in 10827 genes of 15456 genes)", x="Number of Peaks", y="Number of Genes")
```
Look at these genes compared to distribution for number of peaks in all peaks before filter

```{r}
allPAS=read.table("../data/assignedPeaks/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.SAF", header = T, stringsAsFactors = F) %>% separate(GeneID, into = c("peak", "Chrom", "Peakstart", "PeakEnd", "strand", "geneid"),sep=":") %>% separate(geneid, into=c("gene", "loc"), sep="_") %>% group_by(gene) %>% summarise(nPeakAll=n()) %>% mutate(AddedPeak=ifelse(gene %in% uniqnew_genes$gene, "yes", "no"))


ggplot(allPAS, aes(x=nPeakAll))+ geom_histogram(bins = 100) + facet_grid(~AddedPeak)
```

Mean number of peaks for genes where we added, this could help us understand the distribution shift:  

```{r}
allPAS_withnewpas=allPAS %>% filter(AddedPeak=="yes")

summary(allPAS_withnewpas$nPeakAll)

```




Look at n peaks per gene in old data and npeaks per gene in new- join and subtract
```{r}
#newPAS5perc_pergene

oldPAS5perc_pergene= oldPAS5perc%>% separate(ID, into=c("gene", "peaknum"), sep=":") %>% group_by(gene) %>% summarise(nPeakOld=n())

nrow(oldPAS5perc_pergene)
```

Join: 

```{r}
npeakpergenebot=oldPAS5perc_pergene %>% full_join(newPAS5perc_pergene, by="gene") %>% replace_na(list(nPeakOld = 0, nPeak = 0)) %>% mutate(NewMinOld=nPeak-nPeakOld)


ggplot(npeakpergenebot, aes(x=nPeakOld, y=nPeak)) + geom_point() + geom_smooth(method="lm") + annotate("text", label="r2=0.4447", x=12, y=2) + labs(title="Number of Peaks in old vs new data",y="Number of peaks new data", x="Number of peaks old data")
```
Correlation between number of peaks before and after:

```{r}
summary(lm(npeakpergenebot$nPeak~npeakpergenebot$nPeakOld))

```
```{r}
summary(npeakpergenebot$NewMinOld)

ggplot(npeakpergenebot, aes(x=NewMinOld)) + geom_histogram(bins=100)
```


