---
title: "Usage Effect Sizes"
author: "Briana Mittleman"
date: "5/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In order to compare effect sizes for the QTLs I have previously identified in an interpretable manner, I need to run the linear model with the non normalized usage. To do this I will separate the the usage (with annotation) files by chromosome and run fastqtl on these files.  

```{r}
library(workflowr)
library(tidyverse)
library(cowplot)
```

##Prepare files  

countsnum= APApeak_Phenotype_GeneLocAnno.Total.5perc.CountsNumeric, APApeak_Phenotype_GeneLocAnno.Nuclear.5perc.CountsNumeric

id file= APApeak_Phenotype_GeneLocAnno.Nuclear.5perc.fc.gz, APApeak_Phenotype_GeneLocAnno.Total.5perc.fc.gz



```{r}
totAnno= read.table("../data/phenotype_5perc/APApeak_Phenotype_GeneLocAnno.Total.5perc.fc.gz", stringsAsFactors = F, header = T) %>% separate(chrom, into=c("Chrchrom", "Start", "End", "ID"),sep=":") %>% mutate(Chr=str_sub(Chrchrom, 4, str_length(Chrchrom)))
                                                                                                                                                                                                                
colnamesTot= colnames(totAnno)[5:56]
totUsage=read.table("../data/phenotype_5perc/APApeak_Phenotype_GeneLocAnno.Total.5perc.CountsNumeric", stringsAsFactors = F, header = F, col.names = colnamesTot) 

totUsageAnno=as.data.frame(cbind(Chr=totAnno$Chr, start=totAnno$Start, end=totAnno$End, ID=totAnno$ID, totUsage ))

write.table(totUsageAnno,file="../data/nonNorm_pheno/TotalUsageAllChrom.txt", col.names = T, row.names = F, quote = F, sep="\t" )
```



```{r}
nucAnno= read.table("../data/phenotype_5perc/APApeak_Phenotype_GeneLocAnno.Nuclear.5perc.fc.gz", stringsAsFactors = F, header = T)%>% separate(chrom, into=c("Chrchrom", "Start", "End", "ID"),sep=":") %>% mutate(Chr=str_sub(Chrchrom, 4, str_length(Chrchrom)))
colnamesNuc= colnames(nucAnno)[5:56]
nucUsage=read.table("../data/phenotype_5perc/APApeak_Phenotype_GeneLocAnno.Nuclear.5perc.CountsNumeric", stringsAsFactors = F, header = F, col.names = colnamesNuc) 


nucUsageAnno=as.data.frame(cbind(Chr=nucAnno$Chr, start=nucAnno$Start, end=nucAnno$End, ID=nucAnno$ID, nucUsage ))

write.table(nucUsageAnno,file="../data/nonNorm_pheno/NuclearUsageAllChrom.txt", col.names = T, row.names = F, quote = F, sep="\t" )
```


##Run QTL scripts
I will create a python script to seperate the file into each chromosome for running fastQTL.  

```{bash,eval=F}
sbatch run_sepUsagephen.sh
sbatch ZipandTabPheno.sh
sbatch ApaQTL_nominalNonnorm.sh
```

Concatinate files:  

```{bash,eval=F}
cat  ../data/nonNorm_pheno/TotalUsageChrom*.nominal.out > ../data/nonNorm_pheno/TotalUsageChrom_Nominal_AllChrom.txt
cat ../data/nonNorm_pheno/NuclearUsageChrom*.nominal.out > ../data/nonNorm_pheno/NuclearUsageChrom_Nominal_AllChrom.txt

```

##Pull out real total and nuc QLTs  


```{bash,eval=F}
python qtlsPvalOppFrac.py   ../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt ../data/nonNorm_pheno/TotalUsageChrom_Nominal_AllChrom.txt ../data/QTLoverlap_nonNorm/NuclearQTLinTotalNominal_nonNorm.txt  

python qtlsPvalOppFrac.py ../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt ../data/nonNorm_pheno/NuclearUsageChrom_Nominal_AllChrom.txt ../data/QTLoverlap_nonNorm/NuclearQTLinNuclearNominal_nonNorm.txt  


python qtlsPvalOppFrac.py ../data/apaQTLs/Total_apaQTLs4pc_5fdr.txt ../data/nonNorm_pheno/TotalUsageChrom_Nominal_AllChrom.txt ../data/QTLoverlap_nonNorm/TotalQTLinTotalNominal_nonNorm.txt  

python qtlsPvalOppFrac.py ../data/apaQTLs/Total_apaQTLs4pc_5fdr.txt ../data/nonNorm_pheno/NuclearUsageChrom_Nominal_AllChrom.txt ../data/QTLoverlap_nonNorm/TotalQTLinNuclearNominal_nonNorm.txt  
```



```{r}
totAPAinNuc=read.table("../data/QTLoverlap_nonNorm/TotalQTLinNuclearNominal_nonNorm.txt", header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "slope"))


nucAPAinTot=read.table("../data/QTLoverlap_nonNorm/NuclearQTLinTotalNominal_nonNorm.txt", header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "slope"))

totAPAinTot=read.table("../data/QTLoverlap_nonNorm/TotalQTLinTotalNominal_nonNorm.txt", header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "slope")) %>% dplyr::select(peakID, snp, slope) %>% dplyr::rename("Originalslope"=slope)

nucAPAinNuc=read.table("../data/QTLoverlap_nonNorm/NuclearQTLinNuclearNominal_nonNorm.txt", header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "slope")) %>% dplyr::select(peakID, snp, slope)%>% dplyr::rename("Originalslope"=slope)
```



##Total

```{r}
TotBoth= totAPAinNuc %>% inner_join(totAPAinTot,by=c("peakID", "snp"))

summary(lm(TotBoth$slope ~ TotBoth$Originalslope))


totbothplot=ggplot(TotBoth, aes(x=Originalslope, y=slope))+geom_point() + geom_smooth(method="lm") + labs(title="Total apaQTL effect sizes", x="Effect size in Total",y="Effect size in Nucler") + geom_density_2d(col="red") + annotate("text", y=1, x=0, label="R2=.67, slope=0.64")
```
##Nuclear  

```{r}
NucBoth= nucAPAinTot %>% inner_join(nucAPAinNuc,by=c("peakID", "snp"))
summary(lm(NucBoth$slope ~ NucBoth$Originalslope))

Nucbothplot=ggplot(NucBoth, aes(x=Originalslope, y=slope))+geom_point() + geom_smooth(method="lm") + labs(title="Nuclear apaQTL effect sizes", x="Effect size in Nuclear",y="Effect size in Total") + geom_density_2d(col="red") +  annotate("text", y=1, x=0, label="R2=.8, slope=0.78")
```

```{r}
plot_grid(totbothplot,Nucbothplot)
```

For the nuclear plot I want to include the PAS we cannot test in the total fraction. To do this I will write code that only writes out the lines for the PAS not in the total fraction. I need the nominal effect sizes for the nuclear qtls in PAS we could not test in total. First I will get all of the PAS tested in total and figure out the nuclear QTLs not in this set. I will then pull those associations out of the nominal nuclear set. 

```{bash,eval=F}
python nucSpeceffectsize.py
```

```{r}
nucSpec=read.table("../data/QTLoverlap_nonNorm/NuclearSpecQTLinNuclearNominal_nonNorm.txt",header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "Originalslope")) %>% mutate(slope=0,set="Specific") %>% select(peakID, snp, dist, pval, Originalslope, slope,set)
NucBoth_set= NucBoth %>% mutate(set="Both")
NucBothwSpec=bind_rows(NucBoth_set, nucSpec)
```

```{r}
summary(lm(NucBothwSpec$slope ~ NucBothwSpec$Originalslope))

cor.test(y=NucBothwSpec$slope, x=NucBothwSpec$Originalslope)$p.value

nucspecplot=ggplot(NucBothwSpec, aes(x=Originalslope, y=slope))+geom_point(aes( col=set)) + geom_smooth(method="lm") + labs(title="Nuclear apaQTL effect sizes", x="Effect size in Nuclear",y="Effect size in Total") +  annotate("text", y=1, x=0, label="R2=.68, slope=0.66",size=6) + scale_color_brewer(palette = "Dark2")+ theme(text = element_text(size=16), legend.position = "bottom")
```
Do this in the other fraction to see what is looks like.

```{bash,eval=F}
python totSeceffectsize.py
```


```{r}
totSpec=read.table("../data/QTLoverlap_nonNorm/TotalSpecQTLinTotalNominal_nonNorm.txt",header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "Originalslope")) %>% mutate(slope=0,set="Specific") %>% select(peakID, snp, dist, pval, Originalslope, slope,set)
TotBoth_set= TotBoth %>% mutate(set="Both")
TotBothwSpec=bind_rows(TotBoth_set, totSpec)
```

```{r}
summary(lm(TotBothwSpec$slope ~ TotBothwSpec$Originalslope))

totspecplot=ggplot(TotBothwSpec, aes(x=Originalslope, y=slope))+geom_point(aes( col=set)) + geom_smooth(method="lm") + labs(title="Total apaQTL effect sizes", x="Effect size in Total",y="Effect size in Nuclear") +  annotate("text", y=1, x=0, label="R2=.65, slope=0.62", size=6) + scale_color_brewer(palette = "Dark2") + theme(text = element_text(size=16), legend.position = "bottom")

totspecplot
```
```{r}
grideffectplot=plot_grid(nucspecplot,totspecplot)
grideffectplot
```

```{r figure2D, include=FALSE, dev="pdf", fig.height=8, fig.width=20, crop=FALSE}
grideffectplot
```

```{r}

TotBothwSpec %>% filter(set=="Specific") %>% nrow()
NucBothwSpec %>% filter(set=="Specific") %>% nrow()
prop.test(x=c(117,40), n=c(565,378), alternative ="greater")
```

