---
title: "Location of Nuclear Intronic PAS"
author: "Briana Mittleman"
date: "5/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am interested in understanding where in the introns the nuclear peaks are. Are they closer to the three prime or five prime edge of the intron. This may help us understand if NMD is contributing to the loss of transcripts between the nuclear and total fraction.  

I need to create an annotation with introns that do not overlap. For this I will use line up all of the exons for a gene then take the open spaces as introns.  
```{r}

library(tidyverse)
library(workflowr)
library(cowplot)
```


```{r}
nucIntronicPeaks=read.table("../data/peaks_5perc/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.fc", stringsAsFactors = F, header = F,col.names = c("chr", "start", "end", "gene", "loc", "strand", "peak", "avgUsage")) %>% filter(loc=="intron")

nucIntronicPeaksBed=nucIntronicPeaks %>% mutate(ID=paste(peak,gene,loc, sep=":")) %>% dplyr::select(chr, start, end, ID, avgUsage, strand)

write.table(nucIntronicPeaksBed, "../data/intron_analysis/NuclearIntronicPeaks.bed", col.names = F, row.names = F, quote = F,sep="\t")
```

I will need to assign each of these to an intron in the new annotation.  

The genome annotation file, Transcript2GeneName.dms has the information i need. I will need to parse this file. I need all exons for a gene (longest transcript) The file has the exon starts and ends for each transcript.  

I will remove the exon locations for full transcripts using bedtools subtract. 



Create transcript file.I will select all of the transcripts in the dms file and merge by gene name. Then I can subtract the exons 



```{bash,eval=F}
python transcriptdm2bed.py
```

Sort the output, group by transcript and fix order of columns. 

```{bash,eval=F}
sort -k1,1 -k2,2n /project2/gilad/briana/genome_anotation_data/RefSeq_annotations/AllTranscriptsbyName.bed > /project2/gilad/briana/genome_anotation_data/RefSeq_annotations/AllTranscriptsbyName.sort.bed

sbatch grouptranscripts.py

python fixgroupedtranscript.py
```


Exons:

/project2/gilad/briana/genome_anotation_data/RefSeq_annotations/ncbiRefseq_coding.dms  

Exon bed file:/project2/gilad/briana/genome_anotation_data/RefSeq_annotations/refseq.ProteinCoding.bed

```{bash,eval=F}
sbatch subtractExons.sh
```
sort:

```{bash,eval=F}
sort -k1,1 -k2,2n /project2/gilad/briana/apaQTL/data/intron_analysis/transcriptsMinusExons.bed > /project2/gilad/briana/apaQTL/data/intron_analysis/transcriptsMinusExons.sort.bed
```


Next I will map the intronic peaks on these positions. 

```{bash,eval=F}
sbatch assignNucIntonpeak2intronlocs.sh
```


Plot percentage of intron where PAS is.

```{r}
pas2intron=read.table("../data/intron_analysis/IntronPeaksontoIntrons.bed",col.names = c("intronCHR", "intronStart", "intronEnd", "gene", "score", "strand", "peakCHR", "peakStart", "peakEnd", "PeakID", "meanUsage", "peakStrand")) %>% mutate(PASloc=ifelse(strand=="+", peakEnd, peakStart)) %>% dplyr::select(intronStart, intronEnd, gene, strand, PeakID, PASloc ,meanUsage) %>% mutate(intronLength=intronEnd-intronStart , distance2PAS= ifelse(strand=="+", PASloc-intronStart, intronEnd-PASloc), propIntron=distance2PAS/intronLength)



```


```{r}
nuclearplot=ggplot(pas2intron, aes(x=propIntron)) + geom_histogram(bins=50, aes(y=..count../33345)) + labs(title="PAS position within intron \n for Nuclear PAS", y="Proportion of Intronic PAS", x="Proportion of Intronic Length")
```
Facet by usage 0-25, 25-50, 50-75, 75-1
```{r}
pas2intron_usage=pas2intron %>% mutate(UsageCat=ifelse(meanUsage<=.5, "low", "high"))

ggplot(pas2intron_usage, aes(x=propIntron, fill=UsageCat)) + geom_histogram(bins=50, aes(y=..count../33345)) + labs(title="PAS position within intron", y="Proportion of Intronic PAS", x="Proportion of Intronic Length") + facet_grid(~UsageCat)

```


Look at different intron lengths:

First i want to look at the distribution of intorn lengths:

```{r}
ggplot(pas2intron_usage, aes(x=log10(intronLength))) + geom_density()
```

I will look at above and below the mean intron length:

```{r}
meanIntronlength=mean(pas2intron_usage$intronLength)

pas2intron_length=pas2intron %>% mutate(LengthCat=ifelse(intronLength<=meanIntronlength, "bottom", "top"))


ggplot(pas2intron_length, aes(x=propIntron, fill=LengthCat)) + geom_histogram(bins=50, aes(y=..count../33345)) + labs(title="PAS position within intron", y="Proportion of Intronic PAS", x="Proportion of Intronic Length") + facet_grid(~LengthCat)


```
Look at quartiles:  

```{r}
summary(pas2intron_usage$intronLength)

pas2intron_length2=pas2intron %>% mutate(LengthCat=ifelse(intronLength<=4451, "first", ifelse(intronLength>4451 &intronLength<=11142, "second", ifelse(intronLength>11142 &intronLength<=30024, "third", "fourth"))))
                                         
ggplot(pas2intron_length2, aes(x=propIntron, fill=LengthCat)) + geom_histogram(bins=50, aes(y=..count../33345)) + labs(title="PAS position within intron", y="Proportion of Intronic PAS", x="Proportion of Intronic Length") + facet_grid(~LengthCat)



```


Look at distribution in total fraction:  


```{r}
totIntronicPeaks=read.table("../data/peaks_5perc/APApeak_Peaks_GeneLocAnno.Total.5perc.fc", stringsAsFactors = F, header = F,col.names = c("chr", "start", "end", "gene", "loc", "strand", "peak", "avgUsage")) %>% filter(loc=="intron")

totIntronicPeaksBed=totIntronicPeaks %>% mutate(ID=paste(peak,gene,loc, sep=":")) %>% dplyr::select(chr, start, end, ID, avgUsage, strand)

write.table(totIntronicPeaksBed, "../data/intron_analysis/TotalIntronicPeaks.bed", col.names = F, row.names = F, quote = F,sep="\t")
```


map these to the intron file



```{bash,eval=F}
sbatch assignTotIntronpeak2intronlocs.sh
```

```{r}
pas2intronTot=read.table("../data/intron_analysis/TotalIntronPeaksontoIntrons.bed",col.names = c("intronCHR", "intronStart", "intronEnd", "gene", "score", "strand", "peakCHR", "peakStart", "peakEnd", "PeakID", "meanUsage", "peakStrand")) %>% mutate(PASloc=ifelse(strand=="+", peakEnd, peakStart)) %>% dplyr::select(intronStart, intronEnd, gene, strand, PeakID, PASloc ,meanUsage) %>% mutate(intronLength=intronEnd-intronStart , distance2PAS= ifelse(strand=="+", PASloc-intronStart, intronEnd-PASloc), propIntron=distance2PAS/intronLength)

nrow(pas2intronTot)

```

```{r}
totalplot=ggplot(pas2intronTot, aes(x=propIntron)) + geom_histogram(bins=50, aes(y=..count../31954)) + labs(title="PAS position within intron \nfor total PAS", y="Proportion of Intronic PAS", x="Proportion of Intronic Length")
```


```{r}
plot_grid(totalplot, nuclearplot)
```
```{r}
summary(pas2intronTot$intronLength)

pas2intron_totlength2=pas2intronTot %>% mutate(LengthCat=ifelse(intronLength<=4451, "first", ifelse(intronLength>4275 &intronLength<=10682, "second", ifelse(intronLength>10682 &intronLength<=28377, "third", "fourth"))))
                                         
ggplot(pas2intron_totlength2, aes(x=propIntron, fill=LengthCat)) + geom_histogram(bins=50, aes(y=..count../31954)) + labs(title="PAS position within intron", y="Proportion of Intronic PAS", x="Proportion of Intronic Length") + facet_grid(~LengthCat)


```


