---
title: "Nuclear specific apaQTLs"
author: "Briana Mittleman"
date: "6/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to look at nuclear specific apaQTLs. I expect these to be eQTls.

To look at nuclear specific apaQTLs I will test if nuclear QTLs are nominally significant in the total fraction. I created this file in [this analysis](apaQTLoverlap.html) where I tested qtl overlap.  


```{r}
library(tidyverse)
library(workflowr)
```


I test 589 of the apaQTls in the total fraction.
```{r}
nucAPAinTot=read.table("../data/QTLoverlap/NuclearQTLinTotalNominal.txt", header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "slope")) %>% separate(peakID,into=c("chr", "start", "end", "geneID"),sep=":" ) %>% separate(geneID, into=c("gene", "loc", "strand", "peakNum"), sep="_")
nucAPAinTot_NOTSIG=nucAPAinTot %>% filter(pval>.05)
nucAPAinTot_SIG=nucAPAinTot %>% filter(pval<.05)
```

```{r}
hist(nucAPAinTot$pval, breaks=100)
ggplot(nucAPAinTot, aes(x=pval)) + geom_density() + geom_vline(xintercept = .05)
```

I can test if these are more likely to be eGenes. I can get the eGenes by comparing the 

```{r}
explained=read.table("../data/Li_eQTLs/explainedEgenes.txt", header = F, stringsAsFactors = F, col.names = c("gene"))
unexplained=read.table("../data/Li_eQTLs/UnexplainedEgenes.txt", header = F, stringsAsFactors = F, col.names = c("gene"))

allEgenes=as.data.frame(rbind(explained,unexplained))
```


```{r}
nucAPAinTot_NOTSIG_egene=nucAPAinTot_NOTSIG %>% semi_join(allEgenes, by="gene")
nucAPAinTot_SIG_egene=nucAPAinTot_SIG %>% semi_join(allEgenes, by="gene")
```



Proportion nucspec egene

```{r}
nrow(nucAPAinTot_NOTSIG_egene)/nrow(nucAPAinTot_NOTSIG)
nrow(nucAPAinTot_SIG_egene)/nrow(nucAPAinTot_SIG)

```


Difference of proportion test: 

```{r}
prop.test(x=c(nrow(nucAPAinTot_NOTSIG_egene),nrow(nucAPAinTot_SIG_egene)),n=c(nrow(nucAPAinTot_NOTSIG),nrow(nucAPAinTot_SIG)))
```

Write out nuc specific QTLs
```{r}
write.table(nucAPAinTot_NOTSIG,file="../data/QTLoverlap/NucSpecApaQTL.txt", quote=F, col.names = T, row.names = F )
```


Run this the opposite direction:

Total specific:  
```{r}
totAPAinNuc=read.table("../data/QTLoverlap/TotalQTLinNuclearNominal.txt", header = F, stringsAsFactors = F, col.names=c("peakID", "snp", "dist", "pval", "slope")) %>% separate(peakID,into=c("chr", "start", "end", "geneID"),sep=":" ) %>% separate(geneID, into=c("gene", "loc", "strand", "peakNum"), sep="_")
totAPAinNuc_NOTSIG=totAPAinNuc %>% filter(pval>.05)
totAPAinNuc_SIG=totAPAinNuc %>% filter(pval<.05)

```

```{r}
totAPAinNuc_NOTSIG_egene=totAPAinNuc_NOTSIG %>% semi_join(allEgenes, by="gene")
totAPAinNuc_SIG_egene=totAPAinNuc_SIG %>% semi_join(allEgenes, by="gene")
```



Proportion nucspec egene

```{r}
nrow(totAPAinNuc_NOTSIG_egene)/nrow(totAPAinNuc_NOTSIG)
nrow(totAPAinNuc_SIG_egene)/nrow(totAPAinNuc_SIG)

```

```{r}
prop.test(x=c(nrow(totAPAinNuc_NOTSIG_egene),nrow(totAPAinNuc_SIG_egene)),n=c(nrow(totAPAinNuc_NOTSIG),nrow(totAPAinNuc_SIG)))
```
