---
title: "Nuclear specific APA including those not tested in total"
author: "Briana Mittleman"
date: "6/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In my previous analysis found [here](NuclearSpecAPAqtl.html) I took nuclear specific apa QTLs as those tested in total that are not nominally significant in total. In this analysis I will include the nuclear apaQTLs in PAS not tested in total as nuclear specific. These may be important for explaining eQTLs or pQTLs.


```{r}
library(workflowr)
library(tidyverse)
library(cowplot)
```

I will give all of the QTLs an id. 
```{r}
nucQTls=read.table("../data/apaQTLs/Nuclear_apaQTLs4pc_5fdr.txt",header = T, stringsAsFactors = F) %>% mutate(ID=paste(Gene,Peak, sid, sep=":"))
sharedQTLs=read.table("../data/apaQTLs/SharedAPAQTLs.txt", header = T, stringsAsFactors = F) %>% mutate(ID=paste(gene,peakNum, snp, sep=":"))

sharedQTL_ID=as.vector(sharedQTLs$ID)
```

Nuclear Specific:

```{r}
NuclearSpecQTL= nucQTls %>% mutate(Shared=ifelse(ID %in% sharedQTL_ID, "Yes", "No"))
NuclearSpecQTL$Shared=as.factor(NuclearSpecQTL$Shared)
```

I need to input the explained eGenes, unexplained eGenes, and pGenes. For this I will make sure none of the pgenes are eGenes.

```{r}
explained=read.table("../data/Li_eQTLs/explainedEgenes.txt", header = F, stringsAsFactors = F, col.names = c("gene"))
unexplained=read.table("../data/Li_eQTLs/UnexplainedEgenes.txt", header = F, stringsAsFactors = F, col.names = c("gene"))
protein=read.table("../data/Battle_pQTL/psQTLGeneNames.txt",header = F, stringsAsFactors = F,col.names = c("gene"))
```


```{r}
'%!in%' <- function(x,y)!('%in%'(x,y))


protein_only=protein %>% filter(gene %!in% explained$gene & gene %!in% unexplained$gene)

write.table(protein_only, "../data/Battle_pQTL/pQTLGeneNamesONLYP.txt", col.names = F, row.names = F,quote = F, sep="\t")
```


Are nuc specific less likely to be in p genes?

```{r}
NuclearSpecQTL_gene=NuclearSpecQTL %>% mutate(pGene=ifelse(Gene %in% protein_only$gene, "Yes", "No"), uneplained=ifelse(Gene %in% unexplained$gene, "Yes", "No"), explained=ifelse(Gene %in% explained$gene, "Yes","No"))
```

```{r}
nPandShare=nrow(NuclearSpecQTL_gene %>% filter(Shared=="Yes", pGene=="Yes"))/nrow(NuclearSpecQTL_gene)
nPandShare
nPandNotShare=nrow(NuclearSpecQTL_gene %>% filter(Shared=="No", pGene=="Yes"))/nrow(NuclearSpecQTL_gene)
nPandNotShare

```
Only looking at 8 and 2. This isnt very good.  Cant make claim.  

```{r}
nEandShare=nrow(NuclearSpecQTL_gene %>% filter(Shared=="Yes", uneplained=="Yes" |explained=="Yes" ))

nEandShare
nEandNotShare=nrow(NuclearSpecQTL_gene %>% filter(Shared=="No", uneplained=="Yes" |explained=="Yes"))
nEandNotShare


prop.test(x=c(nEandShare,nEandNotShare),n=c(nrow(NuclearSpecQTL_gene),nrow(NuclearSpecQTL_gene)))
```

I want to not count genes with multiple qtl  

```{r}
nGenes=NuclearSpecQTL_gene %>% group_by(Gene) %>% summarise(n=n()) %>% nrow()
nGenes
Egeneandshared=NuclearSpecQTL_gene %>% filter(Shared=="Yes", uneplained=="Yes" |explained=="Yes" ) %>% group_by(Gene) %>% summarise(n=n()) %>% nrow()
Egeneandshared
EgeneandNotshared=NuclearSpecQTL_gene %>% filter(Shared=="No", uneplained=="Yes" |explained=="Yes" ) %>% group_by(Gene) %>% summarise(n=n()) %>% nrow()
EgeneandNotshared


prop.test(x=c(Egeneandshared,EgeneandNotshared),n=c(nGenes,nGenes))
```
This is significant. This means the extra PAS are most likely driving the egene overlap.  

Write these out for other anaylsis.  

```{r}
NuclearSpecQTL_shared= NuclearSpecQTL %>% filter(Shared=="Yes") %>% select(Gene, sid)
write.table(NuclearSpecQTL_shared,file="../data/NucSpeceQTLeffect/SharedApaQTL_nottestinc.txt", col.names = F, row.names = F, sep="\t", quote = F )
NuclearSpecQTL_specific=NuclearSpecQTL %>% filter(Shared=="No")%>% select(Gene, sid)
write.table(NuclearSpecQTL_specific,file="../data/NucSpeceQTLeffect/NucSpecApaQTL_nottestinc.txt", col.names = F, row.names = F, sep="\t", quote = F )
```

