---
title: "apaQTL and PTTqtls"
author: "Briana Mittleman"
date: "7/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am interesting in understanding the PTTqtls a bit more. I want to ask if pttQTLs are apaQTLs. I will ask if the PTTqtl genes are also apaQTL genes.  

```{r}
library(tidyverse)
library(workflowr)
```



PTT
```{r}
totptt=read.table("../data/PrematureQTLPermuted/Total_preterminationPheno.txt.gz.qqnorm_AllChrBH.txt", stringsAsFactors = F, header = T) %>% filter(-log10(bh)>1) %>% separate(pid,into = c("chr", "start", "end", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "Frac"),sep="_") %>% select(gene) %>% unique()

totpttQTL=read.table("../data/PrematureQTLPermuted/Total_preterminationPheno.txt.gz.qqnorm_AllChrBH.txt", stringsAsFactors = F, header = T) %>% filter(-log10(bh)>1)
write.table(totpttQTL, file="../data/pttQTL/TotalPttQTL.txt", col.names = T, row.names = F, quote = F, sep="\t")

nucptt=read.table("../data/PrematureQTLPermuted/Nuclear_preterminationPheno.txt.gz.qqnorm_AllChrBH.txt", stringsAsFactors = F, header = T) %>% filter(-log10(bh)>1) %>% separate(pid,into = c("chr", "start", "end", "geneID"), sep=":") %>% separate(geneID, into=c("gene", "Frac"),sep="_") %>% select(gene) %>% unique()


nucpttQTL=read.table("../data/PrematureQTLPermuted/Nuclear_preterminationPheno.txt.gz.qqnorm_AllChrBH.txt", stringsAsFactors = F, header = T) %>% filter(-log10(bh)>1)
write.table(nucpttQTL, file="../data/pttQTL/NuclearPttQTL.txt", col.names = T, row.names = F, quote = F, sep="\t")
```

APA

```{r}
nucAPA=read.table("../data/apaQTLs/NuclearapaQTLGenes.txt", col.names = c("gene"),stringsAsFactors = F)
totAPA=read.table("../data/apaQTLs/TotalapaQTLGenes.txt", col.names = c("gene"),stringsAsFactors = F)

```

```{r}
totptt %>% inner_join(totAPA,by="gene") %>% nrow()
totptt %>% anti_join(totAPA,by="gene") %>% nrow()


```

```{r}

nucptt %>% inner_join(nucAPA,by="gene") %>% nrow()
nucptt %>% anti_join(nucAPA,by="gene") %>% nrow()


```


There are 8 total ptt genes that are not apaQTL genes and there are 15 nuclear ptt genes that are not apaQTL genes. 


Look at the distribution. To do this I will pull out the ptt associations from the apa data.

I will right a script that takes a the fraction, pulls the ptt associations into 2 files, one for the associations and one for the non associations.  


```{bash,eval=F}
python pttQTLsinapaQTL.py Total
python pttQTLsinapaQTL.py Nuclear
```


```{r}
names=c('pid' ,'nvar' ,'shape1' ,'shape2', 'dummy' ,'sid', 'dist', 'npval', 'slope', 'ppval' ,'bpval', 'bh')
totapaPTT=read.table("../data/pttQTL/Totalapa_TotalPttQTL.txt", col.names = names, stringsAsFactors = F)
totapaNotPTT=read.table("../data/pttQTL/Totalapa_NOT_TotalPttQTL.txt", col.names = names, stringsAsFactors = F)
```

```{r}
qqplot(-log10(runif(nrow(totapaNotPTT))), -log10(totapaNotPTT$bpval),ylab="-log10 expression permuted pvalue", xlab="Uniform expectation", main="Total PTT genes in Total apaQTL")
points(sort(-log10(runif(nrow(totapaPTT)))), sort(-log10(totapaPTT$bpval)),col= alpha("Red"))
abline(0,1)
legend("topleft", legend=c("ptt gene", "All other genes"),col=c("red", "black"), pch=16,bty = 'n')
```

```{r}
nucapaPTT=read.table("../data/pttQTL/Nuclearapa_NuclearPttQTL.txt", col.names = names, stringsAsFactors = F)
nucapaNotPTT=read.table("../data/pttQTL/Nuclearapa_NOT_NuclearPttQTL.txt", col.names = names, stringsAsFactors = F)
```

```{r}
qqplot(-log10(runif(nrow(nucapaNotPTT))), -log10(nucapaNotPTT$bpval),ylab="-log10 expression permuted pvalue", xlab="Uniform expectation", main="Nuclear PTT genes in Nuclear apaQTL")
points(sort(-log10(runif(nrow(nucapaPTT)))), sort(-log10(nucapaPTT$bpval)),col= alpha("Red"))
abline(0,1)

legend("topleft", legend=c("ptt gene", "All other genes"),col=c("red", "black"), pch=16,bty = 'n')
```


These show the sharing is super high and we are getting at similar phenotypes in ptt and apa.  



Is there something interesting about the non overlap genes:  

```{r}
pttOnlyTot=totptt %>% anti_join(totAPA,by="gene")


totapaPTT_pttstatus=totapaPTT %>% separate(pid, into=c("chr", "start", "end", "geneID"),sep=":")  %>% separate(geneID, into=c("gene", "pos", "strand", "pas"),sep="_") %>% mutate(APAQTL=ifelse(gene %in% pttOnlyTot$gene, "No", "Yes"))

```


```{r}
ggplot(totapaPTT_pttstatus,aes(x=APAQTL, y=bpval)) + geom_boxplot()
```
```{r}
pttOnlyTot$gene
```



```{r}
pttOnlyNuc=nucptt %>% anti_join(nucAPA,by="gene")


nucapaPTT_pttstatus=nucapaPTT %>% separate(pid, into=c("chr", "start", "end", "geneID"),sep=":")  %>% separate(geneID, into=c("gene", "pos", "strand", "pas"),sep="_") %>% mutate(APAQTL=ifelse(gene %in% pttOnlyTot$gene, "No", "Yes"))


ggplot(nucapaPTT_pttstatus,aes(x=APAQTL, y=bpval)) + geom_boxplot()
```
```{r}
pttOnlyNuc$gene
```

