---
title: "Recompute usage with old data"
author: "Briana Mittleman"
date: "5/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(reshape2)
library(workflowr)
library(cowplot)
```

In this analysis I will compute the PAS usage for the new phenotypes in the old data. To make the info comparable and I will rerun feature counts with the filtered phenotypes for both the old and the new data.  

Convert the filtered data to an SAF  

```{bash,eval=F}
python finalPASbed2SAF.py ../data/peaks_5perc/APApeak_Peaks_GeneLocAnno.Nuclear.5perc.fc ../data/CompareOldandNew/APApeak_5perc_Nuclear.SAF

python finalPASbed2SAF.py ../data/peaks_5perc/APApeak_Peaks_GeneLocAnno.Total.5perc.fc ../data/CompareOldandNew/APApeak_5perc_Total.SAF
```

Run feature counts:  


```{bash,eval=F}
sbatch FC_newPeaks_olddata.sh 

```


Convert to phenotypes: 

fix headers on FC 
```{bash,eval=F}
python fixFChead.py ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fc ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.fc 

python fixFChead_bothfrac.py ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fc ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.fc

python fixFChead.py ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fc ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.fc 

python fixFChead_bothfrac.py ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fc ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.fc

```


```{bash,eval=F}
python makePheno.py ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.fc ../data/peakCoverage/file_id_mapping_Total_Transcript.txt ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.pheno.fc
python makePheno.py ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.fc ../data/peakCoverage/file_id_mapping_Total_Transcript.txt ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.pheno.fc

python makePheno.py ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.fc  ../data/peakCoverage/file_id_mapping_Nuclear_Transcript.txt ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.pheno.fc 
python makePheno.py ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.fc ../data/peakCoverage/file_id_mapping_Nuclear_Transcript.txt ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.pheno.fc

```

COunts only: 

```{bash,eval=F}
Rscript pheno2countonly.R -I ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.pheno.fc -O ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.pheno.CountsOnly
Rscript pheno2countonly.R -I ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.pheno.fc -O ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.pheno.CountsOnly

Rscript pheno2countonly.R -I ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.pheno.fc -O ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.pheno.CountsOnly
Rscript pheno2countonly.R -I ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.pheno.fc -O ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.pheno.CountsOnly


```



Convert to numeric 

```{bash,eval=F}
python convertNumeric.py ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.pheno.CountsOnly ../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.pheno.CountsOnlyNumeric

python convertNumeric.py ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.pheno.CountsOnly ../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.pheno.CountsOnlyNumeric


python convertNumeric.py ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.pheno.CountsOnly ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.pheno.CountsOnlyNumeric

python convertNumeric.py ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.pheno.CountsOnly ../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.pheno.CountsOnlyNumeric
```


Total New data 
```{r}

totalPeakUs_new=read.table("../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr,-Start,-End,-Strand,-Length)



ind=as.data.frame(colnames(totalPeakUs_new)[2:dim(totalPeakUs_new)[2]])
colnames(ind)=c("x")
ind=ind %>% separate(x,into=c("indiv", "fraction"), sep="_") %>%mutate(Individual=paste("NA",substring(indiv,2, 6), sep=""))


totalPeakUs_new_CountNum=read.table("../data/CompareOldandNew/Total/New5percPeaks_Total_newdata.fixed.pheno.CountsOnlyNumeric", col.names = ind$Individual)

#numeric with anno
totalPeakNew=as.data.frame(cbind(ID=totalPeakUs_new[,1], totalPeakUs_new_CountNum))
```


Total Old data 

```{r}

totalPeakUs_old=read.table("../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr,-Start,-End,-Strand,-Length)



totalPeakUs_old_CountNum=read.table("../data/CompareOldandNew/Total/New5percPeaks_Total_olddata.fixed.pheno.CountsOnlyNumeric", col.names = ind$Individual)

#numeric with anno
totalPeakold=as.data.frame(cbind(ID=totalPeakUs_old[,1], totalPeakUs_old_CountNum))
```
Seperate by batch 

```{r}
batch4=read.table("../data/MetaDataSequencing.txt", header=T,stringsAsFactors = F)%>% filter(fraction=="total") %>%  select(line, batch) %>% filter(batch == 4)
newInd=batch4$line
```

```{r}
totalPeakoldM=melt(totalPeakold, id.vars=c("ID"), value.name = "Usage", variable.name = "Individual") %>% mutate(New15=ifelse(Individual %in%newInd, "Yes", "No")) %>%  group_by(New15,ID) %>% summarise(meanUsageOld=mean(Usage)) %>% spread(New15,meanUsageOld)

totalPeaknewM=melt(totalPeakNew, id.vars=c("ID"), value.name = "Usage", variable.name = "Individual") %>% mutate(New15=ifelse(Individual %in%newInd, "Yes", "No")) %>%  group_by(New15,ID) %>% summarise(meanUsageNew=mean(Usage)) %>% spread(New15,meanUsageNew)




```

```{r}
totalold=ggplot(totalPeakoldM,aes(x=No,y=Yes)) + geom_density2d() + labs(x="39 ind", y="15 ind", title="Total Usage Old data")
totalnew=ggplot(totalPeaknewM,aes(x=No,y=Yes)) + geom_density2d() + labs(x="39 ind", y="15 ind", title="Total Usage New data")
```

```{r}
plot_grid(totalold,totalnew)
```


Nuclear New data 
```{r}

nuclearPeakUs_new=read.table("../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr,-Start,-End,-Strand,-Length)




nuclearPeakUs_new_CountNum=read.table("../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_newdata.fixed.pheno.CountsOnlyNumeric", col.names = ind$Individual)

#numeric with anno
nuclearPeakNew=as.data.frame(cbind(ID=nuclearPeakUs_new[,1],nuclearPeakUs_new_CountNum))
```


Nuclear Old data 

```{r}

nuclearPeakUs_old=read.table("../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr,-Start,-End,-Strand,-Length)



nuclearPeakUs_old_CountNum=read.table("../data/CompareOldandNew/Nuclear/New5percPeaks_Nuclear_olddata.fixed.pheno.CountsOnlyNumeric", col.names = ind$Individual)

#numeric with anno
nuclearPeakold=as.data.frame(cbind(ID=nuclearPeakUs_old[,1], nuclearPeakUs_old_CountNum))
```
Seperate by batch 


```{r}
nuclearPeakoldM=melt(nuclearPeakold, id.vars=c("ID"), value.name = "Usage", variable.name = "Individual") %>% mutate(New15=ifelse(Individual %in%newInd, "Yes", "No")) %>%  group_by(New15,ID) %>% summarise(meanUsageOld=mean(Usage)) %>% spread(New15,meanUsageOld)

nuclearPeaknewM=melt(nuclearPeakNew, id.vars=c("ID"), value.name = "Usage", variable.name = "Individual") %>% mutate(New15=ifelse(Individual %in%newInd, "Yes", "No")) %>%  group_by(New15,ID) %>% summarise(meanUsageNew=mean(Usage)) %>% spread(New15,meanUsageNew)




```

```{r}
nuclearold=ggplot(nuclearPeakoldM,aes(x=No,y=Yes)) + geom_density2d() + labs(x="39 ind", y="15 ind", title="Nuclear Usage Old data")
nuclearnew=ggplot(nuclearPeaknewM,aes(x=No,y=Yes)) +  geom_density2d() + labs(x="39 ind", y="15 ind", title="Nuclear Usage New data")
```

```{r}
plot_grid(nuclearold,nuclearnew)
```


Subset to new peaks:  

```{r}
NewPeak=read.table( file="../data/peaks_5perc/NewVOldPeaks.txt", header = T) %>% filter(New=="new")
```
Subset total:  

```{r}
totalPeakoldM_new=totalPeakoldM %>% separate(ID, into = c("peak", "chr", "Start", "end","strand","id"),sep=":") %>% filter(peak %in% NewPeak$peak)

totalPeaknewM_new=totalPeaknewM %>% separate(ID, into = c("peak", "chr", "Start", "end","strand","id"),sep=":") %>% filter(peak %in% NewPeak$peak)
```

Plot:

```{r}
totaloldnewpeak=ggplot(totalPeakoldM_new,aes(x=No,y=Yes))  + geom_density2d() + labs(x="39 ind", y="15 ind", title="Total Usage Old data \n New Peaks ")
totalnewnewpeak=ggplot(totalPeaknewM_new,aes(x=No,y=Yes))+ geom_density2d() + labs(x="39 ind", y="15 ind", title="Total Usage New data\n New Peaks ")

plot_grid(totaloldnewpeak,totalnewnewpeak)
```

Subset total:  

```{r}
nuclearPeakoldM_new=nuclearPeakoldM %>% separate(ID, into = c("peak", "chr", "Start", "end","strand","id"),sep=":") %>% filter(peak %in% NewPeak$peak)

nuclearPeaknewM_new=nuclearPeaknewM %>% separate(ID, into = c("peak", "chr", "Start", "end","strand","id"),sep=":") %>% filter(peak %in% NewPeak$peak)
```

Plot:

```{r}
nuclearoldnewpeak=ggplot(nuclearPeakoldM_new,aes(x=No,y=Yes)) +  geom_density2d() + labs(x="39 ind", y="15 ind", title="nuclear Usage Old data \n New Peaks ")
nuclearnewnewpeak=ggplot(nuclearPeaknewM_new,aes(x=No,y=Yes)) + geom_density2d() + labs(x="39 ind", y="15 ind", title="nuclear Usage New data\n New Peaks ")

plot_grid(nuclearoldnewpeak,nuclearnewnewpeak)
```
