---
title: "Verify first intron result"
author: "Briana Mittleman"
date: "6/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the [previous analysis](nucintronicanalysis.html) I saw that most of my intronic pas are in the first intron and skew toward the beginning of long introns. I will further explore this result here. 
```{r}
library(tidyverse)
library(cowplot)
library(workflowr)
```


##Nuclear
These are the nuclear intronic PAS 
```{r}
pas2intron=read.table("../data/intron_analysis/IntronPeaksontoIntrons.bed",col.names = c("intronCHR", "intronStart", "intronEnd", "gene", "score", "strand", "peakCHR", "peakStart", "peakEnd", "PeakID", "meanUsage", "peakStrand"),stringsAsFactors = F) %>% mutate(PASloc=ifelse(strand=="+", peakEnd, peakStart)) %>% dplyr::select(intronStart, intronEnd, gene, strand, PeakID, PASloc ,meanUsage) %>% mutate(intronLength=intronEnd-intronStart , distance2PAS= ifelse(strand=="+", PASloc-intronStart, intronEnd-PASloc), propIntron=distance2PAS/intronLength) %>% mutate(LengthCat=ifelse(intronLength<=3929, "first", ifelse(intronLength>3929 &intronLength<=9220, "second", ifelse(intronLength>9220 &intronLength<=24094, "third", "fourth"))))

pas2intron$LengthCat <- factor(pas2intron$LengthCat, levels=c("first", "second", "third", "fourth"))
```


###Beginning of introns
I want to plot the absolute distance rather than the standardized distance to the 5' ss. 


```{r}
ggplot(pas2intron,aes(x=distance2PAS, fill=LengthCat)) + geom_histogram(bins=100)  + facet_grid(~LengthCat) + xlim(0,5000)


ggplot(pas2intron,aes(x=distance2PAS, fill=LengthCat))  + facet_grid(~LengthCat) + xlim(0,5000) +  stat_ecdf(aes(col=LengthCat)) 
```






###First intron 
This is not the correct analysis. I need to actually look at which intron from all of them. 

this is the file I created to get the introns. I need to remove genes with only 1 introm. 

```{r}
introns=read.table("/project2/gilad/briana/apaQTL/data/intron_analysis/transcriptsMinusExons.sort.bed",stringsAsFactors = F, col.names = c("chrom", "intronStart", "intronEnd", "gene", "score", "strand")) %>% group_by(gene)  %>% filter(!grepl("hap",chrom)) %>% mutate(Intronid=ifelse(strand=="+",  1:n(),n():1), nintron=n()) %>% filter(nintron>2)
```

Join with PAS:

```{r}
pas2intron_intron=pas2intron %>% inner_join(introns, by=c("intronStart","intronEnd","gene", "strand" ))
pas2intron_intron$Intronid=as.factor(pas2intron_intron$Intronid)


write.table(pas2intron_intron, "../data/intron_analysis/NuclearIntronPASwithWhichintron.txt", col.names = T, row.names = F, quote = F, sep="\t")
```


```{r}
ggplot(pas2intron_intron,aes(x=Intronid)) +  geom_bar(stat="count") + labs(title="intron ID for nuclear intronic pas", x="intron ID")
```
```{r}
summary(pas2intron_intron$Intronid)
```
I want to see if the usage is the same over this:

```{r}

pas2intron_intron_usagecat= pas2intron_intron %>% mutate(UsageCat=ifelse(meanUsage<=.1, "<.1", ifelse(meanUsage>.1 &meanUsage<=.2, "<.2", ifelse(meanUsage>.2 &meanUsage<=.3, "<.3", ">.3"))))
pas2intron_intron_usagecat$Intronid=as.numeric(as.character(pas2intron_intron_usagecat$Intronid))
```

```{r}
ggplot(pas2intron_intron_usagecat,aes(x=Intronid, fill=UsageCat)) +  geom_bar(stat="count") + labs(title="intron ID for nuclear intronic pas", x="intron ID") + facet_grid(~UsageCat)+ xlim(0,10)
```
Maybe by the number of introns?

```{r}
summary(pas2intron_intron_usagecat$nintron)
```

```{r}
pas2intron_intron_usagecat_introncat= pas2intron_intron_usagecat %>% mutate(IntronCat=ifelse(nintron<=6, "first (<6)", ifelse(nintron>6 &nintron<=11, "second (6-11)", ifelse(nintron>11 &nintron<=18, "third (11-18)", "fourth (>18)"))))

pas2intron_intron_usagecat_introncat$IntronCat <- factor(pas2intron_intron_usagecat_introncat$IntronCat, levels=c("first (<6)", "second (6-11)", "third (11-18)", "fourth (>18)"))
```


```{r}
ggplot(pas2intron_intron_usagecat_introncat,aes(x=Intronid, fill=IntronCat)) +  geom_bar(stat="count") + labs(title="intron ID for nuclear intronic pas", x="intron ID") + facet_grid(~IntronCat) + xlim(0,10)

nuclear_cdf=ggplot(pas2intron_intron_usagecat_introncat,aes(x=Intronid, fill=IntronCat)) +  stat_ecdf(aes(col=IntronCat)) + labs(title="intron ID for Nuclear intronic pas", x="intron ID") + xlim(0,10)+ geom_vline(xintercept = 2) 

```

Number of introns in each and normalize by average intron size.  

```{r}

pas2intron_intron_grouped=pas2intron_intron %>% group_by(Intronid) %>% summarise(nBin=n(), meanSize=mean(intronLength))  %>% mutate(normNBin=nBin/meanSize) 


pas2intron_intron_grouped$Intronid=as.integer(as.character(pas2intron_intron_grouped$Intronid))

ggplot(pas2intron_intron_grouped, aes(x=Intronid, y=normNBin)) +geom_bar(stat="identity") + labs(title="PAS by Intron", y="normalized number in intron category", x="intron category")

#zoom in 1-10  
pas2intron_intron_grouped_small=pas2intron_intron_grouped %>% filter(Intronid <=10)
ggplot(pas2intron_intron_grouped_small, aes(x=Intronid, y=normNBin)) +geom_bar(stat="identity") + labs(title="PAS by Intron", y="normalized number in intron category", x="intron category")
```










##Total  

###First intron   

```{r}
pas2intronTot=read.table("../data/intron_analysis/TotalIntronPeaksontoIntrons.bed",col.names = c("intronCHR", "intronStart", "intronEnd", "gene", "score", "strand", "peakCHR", "peakStart", "peakEnd", "PeakID", "meanUsage", "peakStrand"),stringsAsFactors = F) %>% mutate(PASloc=ifelse(strand=="+", peakEnd, peakStart)) %>% dplyr::select(intronStart, intronEnd, gene, strand, PeakID, PASloc ,meanUsage) %>% mutate(intronLength=intronEnd-intronStart , distance2PAS= ifelse(strand=="+", PASloc-intronStart, intronEnd-PASloc), propIntron=distance2PAS/intronLength) %>%  mutate(LengthCat=ifelse(intronLength<=3785, "first", ifelse(intronLength>3785 &intronLength<=8872, "second", ifelse(intronLength>8872 &intronLength<=22928, "third", "fourth"))))

pas2intronTot$LengthCat <- factor(pas2intronTot$LengthCat, levels=c("first", "second", "third", "fourth"))
pas2intronTot_intron=pas2intronTot %>% inner_join(introns, by=c("intronStart","intronEnd","gene", "strand" ))

write.table(pas2intronTot_intron, "../data/intron_analysis/TotalIntronPASwithWhichintron.txt", col.names = T, row.names = F, quote = F, sep="\t")
```


```{r}
summary(pas2intronTot_intron$nintron)

pas2intronTot_intron_usagecat_introncat= pas2intronTot_intron %>% mutate(IntronCat=ifelse(nintron<=6, "first (<6)", ifelse(nintron>6 &nintron<=11, "second (6-11)", ifelse(nintron>11 &nintron<=18, "third (11-18)", "fourth (>18)"))))

ggplot(pas2intronTot_intron_usagecat_introncat,aes(x=Intronid, fill=IntronCat)) +  geom_bar(stat="count") + labs(title="intron ID for Total intronic pas", x="intron ID") + facet_grid(~IntronCat) + xlim(0,10)



totalcdf=ggplot(pas2intronTot_intron_usagecat_introncat,aes(x=Intronid, fill=IntronCat)) +  stat_ecdf(aes(col=IntronCat)) + labs(title="intron ID for Total intronic pas", x="intron ID") + xlim(0,10) + geom_vline(xintercept = 2) 
```
##Both fracitons  

Plot both:

```{r}
pas2intronTot_intron_usagecat_introncat_frac=pas2intronTot_intron_usagecat_introncat %>% mutate(fraction="Total") %>% select(Intronid,IntronCat,fraction)

pas2intron_intron_usagecat_introncat_frac=pas2intron_intron_usagecat_introncat%>% mutate(fraction="Nuclear") %>% select(Intronid,IntronCat,fraction)

intronidboth=bind_rows(pas2intronTot_intron_usagecat_introncat_frac,pas2intron_intron_usagecat_introncat_frac)
```



```{r}
ggplot(intronidboth,aes(x=Intronid)) +  stat_ecdf(aes(col=fraction)) + labs(title="intron ID for intronic pas", x="intron ID") + xlim(0,10) + facet_grid(~IntronCat)
```

```{r}
plot_grid(nuclear_cdf,totalcdf)
```

Usage in both fractions. 

```{r}
TotalIntronicUsage=pas2intronTot_intron_usagecat_introncat %>% mutate(fraction="Total") %>% select(meanUsage,fraction)

NuclearIntronicUsage=pas2intron_intron_usagecat_introncat%>% mutate(fraction="Nuclear") %>% select(meanUsage,fraction)

bothIntronicUsage=bind_rows(TotalIntronicUsage,NuclearIntronicUsage)
```

```{r}
ggplot(bothIntronicUsage, aes(x=meanUsage))  +  stat_ecdf(aes(col=fraction)) 
```

Final plot:  
first intron (conditioned on the intron being > 2KB) shows no signal (plotting the first 2kb only)

```{r}
firstintron_nuclear=pas2intron_intron_usagecat_introncat %>% filter(Intronid==1,intronLength>2000) 

firstintron_total=pas2intronTot_intron_usagecat_introncat %>% filter(Intronid==1,intronLength>2000) 

```

```{r}
ggplot(firstintron_nuclear,aes(x=distance2PAS, fill=LengthCat)) + geom_histogram(bins=50) +xlim(0,2000) + facet_grid(~LengthCat)+ labs(title="Nuclear intronic PAS in first intron (3025)")

```

```{r}
ggplot(firstintron_total,aes(x=distance2PAS, fill=LengthCat)) + geom_histogram(bins=50) +xlim(0,2000) + facet_grid(~LengthCat) + labs(title="Total intronic PAS in first intron (1804)")
```

