---
title: "Version with 15nt"
author: "Briana Mittleman"
date: "8/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```


```{bash,eval=F}
mkdir ../data/Version15bp6As
mkdir ../data/Version15bp7As
```

I am going to test 2 filtering versions. I will run these in parallel here. I will put each step of the analysis in the directories above. I will start in the SnakefileFiltPAS with the named SAF file. I will need to convert this to a bed file to use bedtools nuc. I will then filter the final SAF and run the quantification. 


These are still on the opposite strand. I will look at the 15 bases upstream of each PAS for T's.

For + strand:
startnew=start-15
endnew=start

for - strand:
startnew=end
endnew=end +15

```{bash,eval=F}
mkdir ../data/Version15bp6As/filter15upfiles
mkdir ../data/Version15bp7As/filter15upfiles
```

##Filter 
```{bash,eval=F}
python SAF215upbed.py ../data/Version15bp6As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15up.bed
python SAF215upbed.py ../data/Version15bp7As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15up.bed
```

Run bedtools nuc on these:  

```{bash,eval=F}
bedtools nuc -s -seq -fi /project2/gilad/briana/genome_anotation_data/genome/Homo_sapiens.GRCh37.75.dna_sm.all.fa -bed ../data/Version15bp6As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15up.bed > ../data/Version15bp6As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC.txt

bedtools nuc -s -seq -fi /project2/gilad/briana/genome_anotation_data/genome/Homo_sapiens.GRCh37.75.dna_sm.all.fa -bed ../data/Version15bp7As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15up.bed > ../data/Version15bp7As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC.txt

```

Filter out 6 or 7 A. I will do this by making a dictionary with the Ok and outputting only the SAF file PAS in this dictionary. 

I will make a script that takes the input, output, the number of A's to filter 

```{bash,eval=F}
python filterSAFforMP.py ../data/Version15bp6As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC.txt ../data/Version15bp6As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC6A.SAF 6

python filterSAFforMP.py ../data/Version15bp7As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC.txt ../data/Version15bp7As/filter15upfiles/APApeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC7A.SAF 7
```

Now i can run feature counts for these files: 

```{bas,eval=F}
mkdir ../data/Version15bp6As/peakCoverage/
mkdir ../data/Version15bp7As/peakCoverage/
sbatch fc_filteredPAS6and7As.sh
```

```{bash,eval=F}
python fixFChead_short.py ../data/Version15bp6As/peakCoverage/APAPeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC6A.Nuclear.Quant.fc ../data/Version15bp6As/peakCoverage/APAPeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC6A.Nuclear.Quant.fixed.fc

python fixFChead_short.py  ../data/Version15bp7As/peakCoverage/APAPeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC7A.Nuclear.Quant.fc ../data/Version15bp7As/peakCoverage/APAPeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC7A.Nuclear.Quant.fixed.fc
```

##Make phenotype 

```{bash,eval=F}
mkdir ../data/Version15bp6As/phenotype/
mkdir ../data/Version15bp7As/phenotype/
  
  
python makePheno.py ../data/Version15bp6As/peakCoverage/APAPeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC6A.Nuclear.Quant.fixed.fc ../data/peakCoverage/file_id_mapping_Nuclear_Transcript.txt  ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.fc


python makePheno.py ../data/Version15bp7As/peakCoverage/APAPeaks.ALLChrom.Filtered.Named.GeneLocAnnoPARSED.15upNUC7A.Nuclear.Quant.fixed.fc ../data/peakCoverage/file_id_mapping_Nuclear_Transcript.txt ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.fc

Rscript pheno2countonly.R -I ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.fc -O ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.CountsOnly

Rscript pheno2countonly.R -I ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.fc -O ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.CountsOnly

python convertNumeric.py ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.CountsOnly ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.CountsOnlyNumeric

python convertNumeric.py ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.CountsOnly ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.CountsOnlyNumeric

mkdir ../data/Version15bp7As/peaks_5perc/
mkdir ../data/Version15bp6As/peaks_5perc/

Rscript filter5perc.R -P ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.fc -N ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.CountsOnlyNumeric -O  ../data/Version15bp6As/peaks_5perc/APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp6Aperc.fc
 
 Rscript filter5perc.R -P ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.fc -N ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.CountsOnlyNumeric -O  ../data/Version15bp7As/peaks_5perc/APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp7Aperc.fc
 
mkdir ../data/Version15bp7As/phenotype_5perc/
mkdir ../data/Version15bp6As/phenotype_5perc/


 python filter5percPheno.py ../data/Version15bp6As/peaks_5perc/APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp6Aperc.fc ../data/Version15bp6As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.fc ../data/Version15bp6As/phenotype_5perc/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.5perc.fc
 
  python filter5percPheno.py ../data/Version15bp7As/peaks_5perc/APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp7Aperc.fc ../data/Version15bp7As/phenotype/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.fc ../data/Version15bp7As/phenotype_5perc/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.5perc.fc
  
  
  #cut -f1-3,7,8,6 -d " " APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp6Aperc.fc > APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp6Aperc.bed
  
 # cut -f1-3,7,8,6 -d " " APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp7Aperc.fc > APApeak_Peaks_GeneLocAnno.Nuclear.5_15bp7Aperc.bed
```

##QTL analysis  


```{bash, eval=F}
module load python

gzip ../data/Version15bp6As/phenotype_5perc/APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.5perc.fc
gzip ../data/Version15bp7As/phenotype_5perc/APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.5perc.fc

#do in dir
python ../../../code/prepare_phenotype_table.py APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.5perc.fc.gz
python ../../../code/prepare_phenotype_table.py APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.5perc.fc.gz
```

```{bash,eval=F}


sh APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.5perc.fc.gz_prepare.sh
head -n 5 APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.5perc.fc.gz.PCs > APApeak_Phenotype_GeneLocAnno_15bp6A.Nuclear.5perc.fc.gz.4PCs


sh APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.5perc.fc.gz_prepare.sh
head -n 5 APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.5perc.fc.gz.PCs > APApeak_Phenotype_GeneLocAnno_15bp7A.Nuclear.5perc.fc.gz.4PCs


mkdir ../data/Version15bp6As/apaQTLPermuted
mkdir ../data/Version15bp6As/apaQTLNominal
mkdir ../data/Version15bp7As/apaQTLPermuted
mkdir ../data/Version15bp7As/apaQTLNominal

sbatch apaQTL_permuted_test6A7A.sh
sbatch apaQTL_nominalv67.sh



cat ../data/Version15bp6As/apaQTLPermuted/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp6A.5perc.fc.gz.qqnorm_chr* > ../data/Version15bp6As/apaQTLPermuted/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp6A.5perc_permRes.txt
cat ../data/Version15bp7As/apaQTLPermuted/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp7A.5perc.fc.gz.qqnorm_chr* > ../data/Version15bp7As/apaQTLPermuted/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp7A.5perc_permRes.txt

cat ../data/Version15bp6As/apaQTLNominal/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp6A.5perc.fc.gz.qqnorm_chr* >../data/Version15bp6As/apaQTLNominal/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp6A.5perc_nomRes.txt
cat ../data/Version15bp7As/apaQTLNominal/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp7A.5perc.fc.gz.qqnorm_chr* >../data/Version15bp7As/apaQTLNominal/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp7A.5perc_nomRes.txt

Rscript apaQTLCorrectedpval_6or7a.R



```

```{r}
QTL6A=read.table("../data/Version15bp6As/apaQTLPermuted/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp6A.5perc_permResBH.txt", stringsAsFactors = F, header = T) %>% separate(pid, into=c("Chr", "Start", "End", "PeakID"), sep=":") %>% separate(PeakID, into=c("Gene", "Loc", "Strand","Peak"), sep="_")

QTL6ASog= QTL6A %>% filter(-log10(bh)>=1)
nrow(QTL6ASog)



```
```{r}
QTL7A=read.table("../data/Version15bp7As/apaQTLPermuted/APApeak_Phenotype_GeneLocAnno.Nuclear_15bp7A.5perc_permResBH.txt", stringsAsFactors = F, header = T) %>% separate(pid, into=c("Chr", "Start", "End", "PeakID"), sep=":") %>% separate(PeakID, into=c("Gene", "Loc", "Strand","Peak"), sep="_")

QTL7Sog= QTL7A %>% filter(-log10(bh)>=1)
nrow(QTL7Sog)
```




##Explained v unexplained QTL  



```{bash,eval=F}

mkdir ../data/Version15bp6As/ApaByEgene
mkdir ../data/Version15bp7As/ApaByEgene

python subsetpermAPAwithGenelist_2versions.py ../data/Li_eQTLs/explainedEgenes.txt 6 ../data/Version15bp6As/ApaByEgene/ApaexplainedeGenes.txt
python subsetpermAPAwithGenelist_2versions.py ../data/Li_eQTLs/explainedEgenes.txt 7 ../data/Version15bp7As/ApaByEgene/ApaexplainedeGenes.txt

python subsetpermAPAwithGenelist_2versions.py ../data/Li_eQTLs/UnexplainedEgenes.txt 6 ../data/Version15bp6As/ApaByEgene/ApaUnaexplainedeGenes.txt
python subsetpermAPAwithGenelist_2versions.py ../data/Li_eQTLs/UnexplainedEgenes.txt 7 ../data/Version15bp7As/ApaByEgene/ApaUnexplainedeGenes.txt


python subsetApanoteGene_2versions.py 6 ../data/Version15bp6As/ApaByEgene/ApaNOTeGene.txt
python subsetApanoteGene_2versions.py 7 ../data/Version15bp7As/ApaByEgene/ApaNOTeGene.txt

```


6As
```{r}
six.notE=read.table("../data/Version15bp6As/ApaByEgene/ApaNOTeGene.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
six.ex=read.table("../data/Version15bp6As/ApaByEgene/ApaexplainedeGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
six.un=read.table("../data/Version15bp6As/ApaByEgene/ApaUnaexplainedeGenes.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval") )
six.un=na.omit(six.un)
```
```{r}
qqplot(-log10(runif(nrow(six.notE))), -log10(six.notE$bpval), xlab="-log10(Uniform)", ylab="-log10(beta pval)", main="Nuclear 6A Apa")
points(sort(-log10(runif(nrow(six.ex)))), sort(-log10(six.ex$bpval)),col= alpha("Red"))
points(sort(-log10(runif(nrow(six.un)))), sort(-log10(six.un$bpval)),col= alpha("Blue"))
abline(0,1)

legend("topleft", legend=c("Not eGenes", "Explained eGenes", "Unexplained eGenes"),col=c("black", "red", "blue"), pch=16,bty = 'n')
```
7As
```{r}
seven.notE=read.table("../data/Version15bp7As/ApaByEgene/ApaNOTeGene.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
seven.ex=read.table("../data/Version15bp7As/ApaByEgene/ApaexplainedeGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
seven.un=read.table("../data/Version15bp7As/ApaByEgene/ApaUnexplainedeGenes.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval") )
seven.un=na.omit(seven.un)
```
```{r}
qqplot(-log10(runif(nrow(seven.notE))), -log10(seven.notE$bpval), xlab="-log10(Uniform)", ylab="-log10(beta pval)", main="Nuclear 7A Apa")
points(sort(-log10(runif(nrow(seven.ex)))), sort(-log10(seven.ex$bpval)),col= alpha("Red"))
points(sort(-log10(runif(nrow(seven.un)))), sort(-log10(seven.un$bpval)),col= alpha("Blue"))
abline(0,1)

legend("topleft", legend=c("Not eGenes", "Explained eGenes", "Unexplained eGenes"),col=c("black", "red", "blue"), pch=16,bty = 'n')
```

```{bash,eval=F}
python convertNominal2SNPloc2Versions.py 6
python convertNominal2SNPloc2Versions.py 7

mkdir ../data/Version15bp6As/overlapeQTL
mkdir ../data/Version15bp7As/overlapeQTL

sbatch run_getapafromeQTL_version6.7.sh
```




##Pqtl  

```{bash,eval=F}
mkdir ../data/Version15bp6As/ApaByPgene
mkdir ../data/Version15bp7As/ApaByPgene

python subsetpermAPAwithGenelist_2versions.py ../data/Battle_pQTL/psQTLGeneNames.txt 6 ../data/Version15bp6As/ApaByPgene/ApaPSGenes.txt

python subsetpermAPAwithGenelist_2versions.py ../data/Battle_pQTL/esQTLGenes.txt 6 ../data/Version15bp6As/ApaByPgene/ApaESGenes.txt

python subsetpermAPAwithGenelist_2versions.py ../data/Battle_pQTL/psQTLGeneNames.txt 7 ../data/Version15bp7As/ApaByPgene/ApaPSGenes.txt

python subsetpermAPAwithGenelist_2versions.py ../data/Battle_pQTL/esQTLGenes.txt 7 ../data/Version15bp7As/ApaByPgene/ApaESGenes.txt

python subsetAPAnotEorPgene_2versions.py 6 ../data/Version15bp6As/ApaByPgene/ApaNOTPorEGenes.txt
python subsetAPAnotEorPgene_2versions.py 7 ../data/Version15bp7As/ApaByPgene/ApaNOTPorEGenes.txt

```

6As

```{r}
six.notEorP=read.table("../data/Version15bp6As/ApaByPgene/ApaNOTPorEGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
six.PS=read.table("../data/Version15bp6As/ApaByPgene/ApaPSGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
six.ES=read.table("../data/Version15bp6As/ApaByPgene/ApaESGenes.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval") )


six_allE=as.data.frame(rbind(six.ex,six.un))
six.PS=na.omit(six.PS)
six.notEorP=na.omit(six.notEorP)
six.ES=na.omit(six.ES)
six.un=na.omit(six.un)
```



```{r}
qqplot(-log10(runif(nrow(six.notEorP))), -log10(six.notEorP$bpval), xlab="-log10(Uniform)", ylab="-log10(beta pval)", main="6A Nuclear Apa")
points(sort(-log10(runif(nrow(six.PS)))), sort(-log10(six.PS$bpval)),col= alpha("Red"))
points(sort(-log10(runif(nrow(six_allE)))), sort(-log10(six_allE$bpval)),col= alpha("blue"))
abline(0,1)

legend("topleft", legend=c("Neither eGenes nor pGenes", "pGenes", "eGenes"),col=c("black", "red","blue"), pch=16,bty = 'n')
```
7As

```{r}
seven.notEorP=read.table("../data/Version15bp7As/ApaByPgene/ApaNOTPorEGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
seven.PS=read.table("../data/Version15bp7As/ApaByPgene/ApaPSGenes.txt",stringsAsFactors = F,col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval"))
seven.ES=read.table("../data/Version15bp7As/ApaByPgene/ApaESGenes.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval") )


seven_allE=as.data.frame(rbind(seven.ex,seven.un))
seven.PS=na.omit(seven.PS)
seven.notEorP=na.omit(seven.notEorP)
seven.ES=na.omit(seven.ES)
seven.un=na.omit(seven.un)
```

```{r}
qqplot(-log10(runif(nrow(seven.notEorP))), -log10(seven.notEorP$bpval), xlab="-log10(Uniform)", ylab="-log10(beta pval)", main="7A Nuclear Apa")
points(sort(-log10(runif(nrow(seven.PS)))), sort(-log10(seven.PS$bpval)),col= alpha("Red"))
points(sort(-log10(runif(nrow(seven_allE)))), sort(-log10(seven_allE$bpval)),col= alpha("blue"))
abline(0,1)

legend("topleft", legend=c("Neither eGenes nor pGenes", "pGenes", "eGenes"),col=c("black", "red","blue"), pch=16,bty = 'n')
```
